### 2022-04-10 paper/paper.tex
Updates from Overleaf

%% Author_tex.tex
%% V1.0
%% 2012/13/12
%% developed by Techset
%%
%% This file describes the coding for rstrans.cls

\documentclass[openacc]{rstransa}%%%%where rstrans is the template name

%%%% *** Do not adjust lengths that control margins, column widths, etc. ***

%%%%%%%%%%% Defining Enunciations  %%%%%%%%%%%
\newtheorem{theorem}{\bf Theorem}[section]
\newtheorem{condition}{\bf Condition}[section]
\newtheorem{corollary}{\bf Corollary}[section]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\Reyn}{\mathrm{Re}}

\begin{document}

%%%% Article title to be placed here
\title{Generalized Quasilinear simulations of Taylor-Couette Flow}

\author{%%%% Author details
Jeffrey S. Oishi$^{1}$ and Morgan Baxter$^{1,2}$}

%%%%%%%%% Insert author address here
\address{$^{1}$Department of Physics and Astronomy, Bates College, Lewiston, ME USA\\
$^{2}$Company Name, Whereever, USA}

%%%% Subject entries to be placed here %%%%
\subject{xxxxx, xxxxx, xxxx}

%%%% Keyword entries to be placed here %%%%
\keywords{xxxx, xxxx, xxxx}

%%%% Insert corresponding author and its email address}
\corres{Jeffrey S. Oishi\\
\email{joishi@bates.edu}}

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
To do: write abstract
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%% Insert the texts which can accomdate on firstpage in the tag "fmtext" %%%%%

\begin{fmtext}
\section{Introduction}
Something about Taylor (1923); GQL.

\section{Equations}

Sample equations.

\end{fmtext}

%%%%%%%%%%%%%%% End of first page %%%%%%%%%%%%%%%%%%%%%

\maketitle

\section{Runs}
Taylor-Couette flow is characterized by $\Reyn_i$, $\Reyn_o$, $\eta$, $\Gamma$. For all runs in this paper, the outer cylinder is stationary, so $\Reyn_o = 0$.

The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}

\begin{table}[!h]
\caption{Simulations}%%%Table caption goes here
\label{tab:simulations}
\begin{tabular}{rllllllll}%%%The number of columns has to be defined here
\hline
Run & $\Reyn_i$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$ & $\Lambda_z$\\
\hline
A & 750 & 0.875 & 3 & 32 & 64 & 64 & - & -\\
B & 1000 & 0.875 & 3 & 32 & 128 & 64 & - & -\\
C & 1500 & 0.875 & 3 & 32 & 128 & 64 & - & -\\
\end{tabular}
\vspace*{-4pt}
\end{table}%%%End of the table

\section{Figures \& Tables}

The output for figure is:\vspace*{-7pt}

%\begin{figure}[!h]
%\centering\includegraphics[width=2.5in]{xxxxxx.eps}
%%% where xxxxxx name represents "figurename.eps"
%\caption{Insert figure caption here}
%\label{fig_sim}
%\end{figure}

\section{Conclusion}

\enlargethispage{20pt}

%\ethics{Insert ethics statement here if applicable.}

\dataccess{Insert details of how to access any supporting data here.}

\aucontribute{JSO designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script. MB performed the simulations, developed data analysis techniques, and added the GQL terms to the TC script. All authors drafted the manuscript.}

\competing{The authors declare that they have no competing interests.}

\funding{Who paid Morgan? Who paid me?}

\ack{We thank Ben Brown for his help validating the Taylor-Couette code. Computations for this paper were performed on the \emph{Leavitt} cluster at the Bates High Performance Computing Center.}

%%%%%%%%%% Insert bibliography here %%%%%%%%%%%%%%

\bibliographystyle{rsta}
\bibliography{TC}

\end{document}

### 2022-04-11 python/taylor_couette_3d.py
Updated TC initial conditions.

@@ -35,6 +35,18 @@ from mpi4py import MPI
 from GQLProjection import GQLProjection
 from filter_field import filter_field
 
+def cond_number(solver):
+    cond_nums = np.zeros(len(solver.pencils))
+    for i,p in enumerate(solver.pencils):
+        cond_nums[i] = np.linalg.cond(p.L.todense())
+    comm = solver.domain.dist.comm
+    global_max = np.array([0.,])
+    local_max = np.array([cond_nums.max(),])
+    comm.Reduce(local_max,global_max, MPI.MAX)
+    if comm.rank == 0:
+        logger.info("Maximum condition number is {:e}".format(global_max[0]))
+
+
 logger = logging.getLogger(__name__)
 
 de.operators.parseables["Project"] = GQLProjection
@@ -92,9 +104,9 @@ Default parameters from Barenghi (1991, J. Comp. Phys.).
 #Lz = 2.0074074463832545
 Sc = 1
 dealias = 3/2
-nz = 32
-ntheta = 64 #32
-nr = 32
+nz=64
+ntheta=128
+nr=32
 
 #eta_string = "{:.4e}".format(eta).replace(".","-")
 
@@ -233,7 +245,7 @@ if GQL:
     problem.substitutions['w_h'] = "Project_high(w)"
     
     # r momentum (GQL)
-    problem.add_equation("r*r*dt(u) - nu*Lap_r - 2*r*v0*v + r*v0*dtheta(u) + r*r*dr(p) = r*v0*v0 - Project_high(r*r*u_h*dr(u_l)+ r*v_h*dtheta(u_l) + r*r*w_h*dz(u_l) - r*v_h*v_l + r*r*u_l*dr(u_h) +r*v_l*dtheta(u_h) + r*r*w_l*dz(u_h) - r*v_h*v_l) - Project_low(r*r*u_h*dr(u_h)+ r*v_h*dtheta(u_h) + r*r*w_h*dz(u_h) - r*v_h*v_h + r*r*u_l*dr(u_l) + r*v_l*dtheta(u_l) + r*r*w_l*dz(u_l) - r*v_l*v_l)")
[…]

### 2022-04-11 python/taylor_couette_3d.py
Added back code to get u from A...

@@ -353,6 +353,16 @@ else:
         A['g']
         A.set_scales(domain.dealias,keep_data=True)
         A['g'] *= fr
+    for vel in [u, v, w]:
+        vel.set_scales(domain.dealias, keep_data=True)
+    # Curl of A
+    u['g'] = A0 * (Az.differentiate('theta')['g']/r - Atheta.differentiate('z')['g'])
+    v['g'] = A0 * (Ar.differentiate('z')['g'] - Az.differentiate('r')['g'])
+    w['g'] = A0 * (Atheta['g'] + r*Atheta.differentiate('r')['g'] - Ar.differentiate('theta')['g'])/r
+    u.differentiate('r', out=ur)
+    v.differentiate('r', out=vr)
+    w.differentiate('r', out=wr)
+
 solver.stop_iteration = np.inf #2000
 
 #CFL stuff

### 2022-04-11 python/taylor_couette_3d.py
Restored missing stop criteria; added final div(u) fix

@@ -358,11 +358,16 @@ else:
     # Curl of A
     u['g'] = A0 * (Az.differentiate('theta')['g']/r - Atheta.differentiate('z')['g'])
     v['g'] = A0 * (Ar.differentiate('z')['g'] - Az.differentiate('r')['g'])
-    w['g'] = A0 * (Atheta['g'] + r*Atheta.differentiate('r')['g'] - Ar.differentiate('theta')['g'])/r
+    w['g'] = A0 * (Atheta['g']/r + Atheta.differentiate('r')['g'] - Ar.differentiate('theta')['g']/r)
     u.differentiate('r', out=ur)
     v.differentiate('r', out=vr)
     w.differentiate('r', out=wr)
 
+#Setting Simulation Runtime
+omega1 = 1/eta - 1.
+period = 2*np.pi/omega1
+solver.stop_sim_time = 10 * period # np.inf #6*period
+solver.stop_wall_time = 24 * 3600. #np.inf # This is in seconds
 solver.stop_iteration = np.inf #2000
 
 #CFL stuff

### 2022-07-28 python/plot_snapshots_tz.py
finalized theta-z snapshot plotting.

@@ -14,9 +14,12 @@ import numpy as np
 import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
+from matplotlib import ticker
+from matplotlib import transforms
 plt.ioff()
 plt.style.use('prl')
-
+import logging
+logger = logging.getLogger(__name__)
 from dedalus.extras import plot_tools
 
 
@@ -31,13 +34,18 @@ def main(filename, start, count, output, period):
             if "_r_slice" in task:
                 tasks.append(task)
             
-    scale = 4
-    dpi = 100
-    title_func = lambda sim_time: r'$t/T_i = {:.3f}$'.format(sim_time)
-    savename_func = lambda write: 'write_tz_{:06}.png'.format(write)
+    scale = 2
+    dpi = 300
+    eta = 0.883
+    Gamma = 29.9
+    Rmid = eta/(1-eta) + 0.5
+    Lx = Rmid*2*np.pi
+    Lz = Gamma
+    title_func = lambda sim_time: r'$t/\tau_\nu = {:.3f}$'.format(sim_time)
+    savename_func = lambda write: 'urms_tz_{:06}.png'.format(write)
     # Layout
-    nrows, ncols = 3,1 
-    image = plot_tools.Box(6, 3)
+    nrows, ncols = 1,1
+    image = plot_tools.Box(3*Lx/Lz, 3)
     pad = plot_tools.Frame(0.2, 0.2, 0.1, 0.1)
     margin = plot_tools.Frame(0.3, 0.2, 0.1, 0.1)
 
@@ -47,19 +55,46 @@ def main(filename, start, count, output, period):
     # Plot writes
     with h5py.File(filename, mode='r') as file:
         for index in range(start, start+count):
-            for n, task in enumerate(tasks):
-                # Build subfigure axes
-                i, j = divmod(n, ncols)
-                axes = mfig.add_axes(i, j, [0, 0, 1, 1])
-                # Call 3D plotting helper, slicing in time
-                dset = file['tasks'][task]
-                image_axes = [2, 1]
-                data_slices = [index, slice(None), slice(None),0]
-                plot_tools.plot_bot(dset, image_axes, data_slices, axes=axes, title=task, even_scale=True)
+            logger.info(f"plotting index {index}")
+            axes = mfig.add_axes(0, 0, [0, 0, 1, 1])
+            # Call 3D plotting helper, slicing in time
[…]

### 2022-07-28 paper/paper.tex
Updates from Overleaf

@@
 \newcommand{\Reyn}{\mathrm{Re}}
-
+\newcommand{\Lop}{\mathcal{L}}
 \begin{document}
@@
 %%%% Article title to be placed here
-\title{Generalized Quasilinear simulations of Taylor-Couette Flow}
+\title{Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Flow}
 
@@
 \begin{abstract}
-To do: write abstract
+
 \end{abstract}
@@
 \begin{fmtext}
-\section{Introduction}
-Something about Taylor (1923); GQL.
-
-\section{Equations}
-
-Sample equations.
-
 \end{fmtext}
-
 %%%%%%%%%%%%%%% End of first page %%%%%%%%%%%%%%%%%%%%%
@@
 
+\section{Introduction}
+\label{sec:intro}
+%Something about Taylor (1923); GQL.
+Since Taylor's classic 1923 paper\cite{1923RSPTA.223..289T}, fluid between two independently rotating cylinders has become a rich laboratory for the understanding of instability, transition to turbulence, and turbulence itself.
+Indeed, the interplay between experiment, computation, and theoretical understanding evinced in Taylor's paper remains a model for the study of fluid dynamics.
+
+Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects.
+In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that contribute to the mean flow.
+Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes.
+GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows \cite{2016PhRvL.116u4501M,2017JFM...810..412T, 2022arXiv220505513M}.
+
+Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$.
+This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the typical cascade found at high $\Reyn$.
+%GQL has been show to outperform QL models in numerous highly turbulent flows, driven by $\beta$-effects on rotating spheres, rotating convection, rotating plane Couette flow (pCf, here after). 
+One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
+In the linearly stable regions, transient growth has been observed in TC flow.
+As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow can be tuned to control the degree of non-normality, that is its commutator with its adjoint
+\begin{equation}
+    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn, \Gamma)
+\end{equation}
+can be zero or non-zero. 
+This provides a very useful arena to test the validity of GQL in a system that admits both normal and non-normal operators.
+
+We pursue the flow into the \emph{spiral turbulence} flow regime, and consider the effectiveness of GQL in each case.
+The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a subcritical transition to spiral turbulence when the cylinders were counter-rotating.
+More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}.
+Using these prior works as a guide, we consider three cases
+\begin{enumerate}
+    \item Inner cylinder rotating only $\Reyn_o = 0$
+    \item Counter rotating cylinders with moderate outer cylinder rotation $\Reyn_o = -1339$
+    \item Counter rotating cylinders with high outer cylinder rotation $\Reyn_o = -3389$
+\end{enumerate}
+Case 1 has a purely normal linear operator, while the latter two feature moderate and significant non-normality respectively.
+
+We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalization, and numerical methods.
+Section~\ref{sec:linear} reviews both modal and non-modal linear analyses and contains a new analysis combining linear and non-linear aspects.
+In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL in each of the three regimes.
+Finally, we offer discussion and concluding results in section~\ref{sec:conclusion}.
+%Recently, a subcritical transition to turbulence via interpenetrating spirals has been reported \cite{2020JFM...892A..12C}.
+This work has a small aspect ratio $\Gamma = 5.26$, which delays the onset of linear instability to considerably higher $\Reyn_1$ for a given $\Reyn_2$.
+
+\section{Methods}
+\label{sec:methods}
+We solve the incompressible Navier-Stokes equations with unit density for $\mathbf{u}' = \mathbf{u} - \mathbf{U}$,
+where $\mathbf{U} = A r + B/r \hat{\mathbf{e}}_\phi$ is circular Couette flow (CCF here after),
+\begin{align}
+      \frac{\partial \mathbf{u}}{\partial t} + \mathbf{u} \cdot \nabla \mathbf{u} &= -\nabla p + \nu \nabla^2\mathbf{u}\\
+      \nabla \cdot \mathbf{u} &= 0.
+\end{align}
+
+We use the Dedalus framework \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
+
+We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_1 = R_1 \Omega_1$.
+Taylor-Couette flow can be parameterised in terms of four parameters, traditionally the inner cylinder Reynolds number $\Reyn_1$, the ratio of inner to outer radii $\eta \equiv R_1/R_2$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_2/\Omega_1$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$.
+In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length.
+We will also refer to the two Reynolds numbers $\Reyn_1$, $\Reyn_2$, related by
+\begin{equation}
+    \frac{\Reyn_2}{\Reyn_1} = \frac{\mu}{\eta}.
+\end{equation}
+
+It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_1$.F
+\section{Linear Stability}
+\label{sec:linear}
+\subsection{Non-Axisymmetric Instability with $\mu = 0$}
+%HAHAH This is all laid out in Krueger, Gross, Di Prima (1966), p 523-524
+%The axisymmetric test case of \cite{1984JFM...146...65M} has $\mu = 0$, $\eta = 0.875$, $\Reyn = 139.31064$, and %$k_{min} = 2\pi/L_z = 2.513$ ($L_z = 2.5$).
+Interestingly, table~\ref{tab:critical_m} shows that circular Couette %flow is linearly unstable to all $m \leq 4$ modes at this $\Reyn$.
+%
+%A natural assumption is that one would not expect to see axisymmetric flow at these parameters.
+Fully non-linear %simulations initialized from very low-amplitude random, $\mathbf{\nabla \cdot u} = 0$ perturbations bear this out.
+%Yet, \cite{1984JFM...146...65M} makes clear reference to TVF being stable to perturbations at this flow.
+He also %claims that initial conditions including ``round-off noise of the computer'' lead to axisymmetric flow.
+This is %quite unlikely to be true.
+%
+%% the two runs (axisymmetric TVF and perturbed, restarted run) are on Leavitt
+%One is naturally led to the question of how axisymmetric TVF is seen in experiment if $\Reyn_{c,m=1}/\Reyn_{c, %m=0} \simeq 1.008$: $m=1$ modes are unstable at a supercriticality of less than 1\%! 
+%The answer becomes obvious if one considers the usual experimental procedure: spin up a fluid from rest to a %$\Reyn < \Reyn_{c,m=0}$, allow Couette flow to be established, then spin up until the first bifurcation occurs.
+%Allow the TVF to settle and proceed upwards.
+This will indeed produce an axisymmetric TVF flow that is %\emph{itself stable against $m\ne 0$ perturbations}!
+At $\Reyn = 138.31$, axisymmetric TVF flow can be perturbed %with gaussian noise and will remain axisymmetric.
+%
+%\begin{table}[!h]
+%\caption{Critical inner $\Reyn_i$ and $k_z$ as a function of $m$ for $\eta = 0.875$, $\mu = 0$}%%%Table caption %goes here
+%\label{tab:critical_m}
+%\begin{tabular}{rll}%%%The number of columns has to be defined here
+%\hline
+%m & $\Reyn_{ic}$ & $k_z$ \\
+%\hline
+%0 & 118.1 & 3.12947\\
+%1 & 119.0 & 3.14374\\
+%2 & 121.5 & 3.18888\\
+%3 & 126.17 & 3.25852\\
+%4 & 133.32 & 3.35672\\
+%5 & 144.20 & 3.47885\\
+%\end{tabular}
+%\vspace*{-4pt}
+%\end{table}%%%End of the table
+
+The axisymmetric test case of \cite{1984JFM...146...65M} has $\mu = 0$, $\eta = 0.875$, $\Reyn = 139.31064$, and $k_{min} = 2\pi/L_z = 2.513$ ($L_z = 2.5$).
+Interestingly, circular Couette flow is linearly unstable to all $m \leq 4$ modes at this $\Reyn$.
+In fact, linear stability analysis shows that $\Reyn_{c,m=1}/\Reyn_{c, m=0} \simeq 1.008$: $m=1$ modes are unstable at a supercriticality of less than 1\%. 
+One is naturally led to the question of how axisymmetric TVF is seen in experiment!
+The answer becomes clear if one considers the usual experimental procedure: spin up a fluid from rest to a $\Reyn < \Reyn_{c,m=0}$, allow Couette flow to be established, then spin up until the first bifurcation occurs.
+Allow the TVF to settle and proceed upwards.
+This will indeed produce an axisymmetric TVF flow that is \emph{itself stable against $m\ne 0$ perturbations}!
+At $\Reyn = 138.31$, axisymmetric TVF flow can be perturbed with gaussian noise and will remain axisymmetric.
+
+This is important for understanding the bifurcations from TVF to WVF: starting from $\Reyn < \Reyn_{W}$ with random initial conditions, one will find a non-axisymmetric flow.
+
+\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
+Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
+\begin{itemize}
+    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
+    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
+    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
+    \item Compare GQL performance at various cutoffs
+    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
+\end{itemize}
+\subsection{Spiral Turbulence}
+Meseguer et al (2009 PRE): 
+\begin{itemize}
+    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
+    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
+\end{itemize}
+
+The spiral turbulence regime with $\mu < 0$ provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_1$ and $\Reyn_2$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.
+
 \section{Runs}
+\label{sec:nonlinear}
 Taylor-Couette flow is characterized by $\Reyn_i$, $\Reyn_o$, $\eta$, $\Gamma$. For all runs in this paper, the outer cylinder is stationary, so $\Reyn_o = 0$.
 
-The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}
+The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}, particularly section 4 of \cite{1984JFM...146...65M}.
+
+First, consider a set of runs with $\mu = 0$.
+This is a standard, linearly unstable case.
+Here, we choose three $\Reyn$ values, one in the WVF flow regime $\Reyn =750$, one in the MWF $\Reyn = 1500$, and one in the TTV regime $\Reyn = 2000$.
+We perform two sets of runs, each varying $\Lambda = {0, 1, 2, 6, 10, 20}$ for both the $z$ and $\theta$ directions.
+The first set of runs is a cold start from identical initial conditions.
+This allows us to understand the ability of GQL to find certain flow states.
+The second set of runs will
 
 \begin{table}[!h]
-\caption{Simulations}%%%Table caption goes here
+\caption{Simulations}
 \label{tab:simulations}
-\begin{tabular}{rllllllll}%%%The number of columns has to be defined here
+\begin{tabular}{rlllllllll}
 \hline
-Run & $\Reyn_i$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$ & $\Lambda_z$\\
+Run & $\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
 \hline
-A & 750 & 0.875 & 3 & 32 & 64 & 64 & - & -\\
-B & 1000 & 0.875 & 3 & 32 & 128 & 64 & - & -\\
-C & 1500 & 0.875 & 3 & 32 & 128 & 64 & - & -\\
+A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
+B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
+C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
+\hline
+D & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
+E & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Subcritical SPT\\
+\hline
+F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
 \end{tabular}
@@
 
-\section{Figures \& Tables}
+\section{Resolution}
+Might compare poorly resolved DNS to low order GQL as done in\cite{2017JFM...810..412T}. Note that in poorly resolved TC DNS, one gets spurious spiral modes.
 
-The output for figure is:\vspace*{-7pt}
+%\section{Conjugate Stable Modes}
+%Conjugate stable modes?
+Typically, one considers the effect of the inviscid stable mode rather than the viscously damped mode.
+However, in the case of TC flow, the background flow profile could not be maintained by the cylinder walls in the limit $\nu \to 0$.
+However, because circular Couette flow is independent of $\Reyn$, probably the limit is OK?
+Can think about this.
+Obviously Rayleigh criterion applies in inviscid case, and this can be easily checked for any angular momentum profile $r^2 \Omega(r)$, so there is some merit to thinking about ideal case even in TC flow.
+Need to think more about this/run it by Adrian.
+
+%The output for figure is:\vspace*{-7pt}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 
+\appendix{Odds and ends}
+\begin{itemize}
+    \item On the $\Reyn_c(m=1)/\Reyn_c(m=0) \simeq 1.008$ issue...there's a discussion in  Krueger, Gross, Di Prima (1966), p 523-524, but it references a solution in a conference proceeding that is difficult to find.
+I have some memory of it being discussed in a more recent paper, but I can't find it anywhere...
+\end{itemize}
 \enlargethispage{20pt}
@@
 
-\aucontribute{JSO designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script.
-MB performed the simulations, developed data analysis techniques, and added the GQL terms to the TC script.
-All authors drafted the manuscript.}
+\aucontribute{JSO wrote the manuscript, designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script.
+MB added the GQL terms to the TC script, developed data analysis techniques, and performed preliminary simulations.
+Both authors edited the manuscript.}
 
@@
 
-\ack{We thank Ben Brown for his help validating the Taylor-Couette code. Computations for this paper were performed on the \emph{Leavitt} cluster at the Bates High Performance Computing Center.}
+\ack{We thank Ben Brown for his help validating the Taylor-Couette code and John Mieszczanski for helping with resolution studies.
+Computations for this paper were performed on the \emph{Leavitt} cluster at the Bates High Performance Computing Center and the NASA Pleiades system under allocation s2276.}
 
@@
 
-\bibliographystyle{rsta}
+\bibliographystyle{RS}
 \bibliography{TC}

### 2022-07-29 python/LSB.py
Added LSB figure and code.

@@ -1,0 +1,83 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+plt.style.use('prl')
+andereck = np.loadtxt("data/andereck_lsb.csv",delimiter=',')
+ours     = np.loadtxt("data/our_LSB.csv",delimiter=',')
+our_path=  np.loadtxt('data/our_path_1.csv',delimiter=',')
+our_path_2=np.loadtxt('data/our_path_2.csv',delimiter=',')
+
+# conversion factors for Coles (1965)
+# ri_c = 5.
+# ro_c = 11.44/2
+# d_c = ro_c - ri_c
+# ifactor = d_c/ri_c
+# ofactor = d_c/ro_c
+
+# coles_H1z = np.loadtxt("data/coles_1965_zoom.csv",delimiter=',')
+# cz_x = ofactor*(coles_H1z[:,2] + coles_H1z[:,0])/2
+# cz_y= ifactor*(coles_H1z[:,3] + coles_H1z[:,1])/2
+# cz_xerr = ofactor*(coles_H1z[:,2] - coles_H1z[:,0])/2
+# cz_yerr = ifactor*(coles_H1z[:,3] - coles_H1z[:,1])/2
+
+# andereck = np.loadtxt("data/andereck_H2.csv",delimiter=',')
+# andereck_x    = (andereck[:,2] + andereck[:,0])/2
+# andereck_y    = (andereck[:,3] + andereck[:,1])/2
+# andereck_xerr = (andereck[:,2] - andereck[:,0])/2
+# andereck_yerr = (andereck[:,3] - andereck[:,1])/2
+
+# coles_H1 = np.loadtxt("data/coles_1965_H1.csv",delimiter=',')
+# c_xerr = ofactor*(coles_H1[:,2] - coles_H1[:,0])
+# c_yerr = ifactor*(coles_H1[:,3] - coles_H1[:,1])
+
+# meseguer_H1 = np.loadtxt("data/meseguer_H1.csv",delimiter=',')
+# meseguer_H2 = np.loadtxt("data/meseguer_H2.csv",delimiter=',')
+# deguchi_H1 = np.loadtxt("data/deguchi_etal_coles.csv", delimiter=',')
+# d_xerr = deguchi_H1[:,2] - deguchi_H1[:,0]
+# d_yerr = deguchi_H1[:,3] - deguchi_H1[:,1]
[…]

### 2022-07-30 python/paths.py
more data analysis

@@ -9,3 +9,17 @@ path2 = ["TC_3d_re_6.400000e+02_eta_8.830000e-01_mu_-1.875000e+00_Gamma_2.990000
          "TC_3d_re_5.400000e+02_eta_8.830000e-01_mu_-2.222222e+00_Gamma_2.990000e+01_M1_5_512_512_64",
          "TC_3d_re_5.200000e+02_eta_8.830000e-01_mu_-2.307692e+00_Gamma_2.990000e+01_M1_5_512_512_64"
          ]
+
+
+rei900_reo_3398 = ["TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64/",
+                   "TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_0_Lambdat_0/",
+                   "TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_1_Lambdat_1/",
+                   "TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_3_Lambdat_3/",
+                   "TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_5_Lambdat_5/",
+                   "TC_3d_re_9.000000e+02_eta_8.830000e-01_mu_-3.333333e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_10_Lambdat_10/",
+]
+
+rei640_reo_1359 = ["TC_3d_re_6.400000e+02_eta_8.830000e-01_mu_-1.875000e+00_Gamma_2.990000e+01_M1_5_512_512_64",
+                   "TC_3d_re_6.400000e+02_eta_8.830000e-01_mu_-1.875000e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_0_Lambdat_0",
[…]

### 2022-07-30 python/plot_KE_DNS_GQL.py

@@ -1,0 +1,34 @@
+from pathlib import Path
+import h5py
+from series import RunSeries
+import matplotlib.pyplot as plt
+import numpy as np
+#from paths import rei640_reo_1359
+from paths import rei900_reo_3398
+
+plt.style.use('prl')
+#path = RunSeries(rei640_reo_1359, 'rei640_reo_-1359')
+path = RunSeries(rei900_reo_3398, 'rei900_reo_-3398')
+
+filetype = Path("scalar/scalar.h5")
+path.data = []
+path.t = []
+t_visc = path.Re[0]
+print(f"t_visc = {t_visc}")
+for i, r in enumerate(path.run_list):
+    filename = path.root_dir/Path(r)/filetype
+    with h5py.File(filename, "r") as df:
+        path.data.append(df['tasks/pertubation_KE'][:,0,0,0])
+        path.t.append(df['scales/sim_time'][:])
+    label = path.Lambdaz[i]
+
+    if label is None:
+        label = '\infty'
+    print(f"Lambda = {label} number of samples = {path.data[i].shape[0]}")
+    plt.plot(path.t[i]/t_visc, path.data[i], label=f"$\Lambda = {label}$")
+plt.xlabel(r"$t/\tau_\nu$")
+plt.ylabel(r"$E_{kin}$")
+plt.xlim(1,2.5)
+plt.legend()
+plt.tight_layout()
+plt.savefig(f"../figs/{path.name}_KE_vs_t.pdf")

### 2022-07-30 python/plot_v_r_DNS_GQL.py

@@ -1,0 +1,31 @@
+from pathlib import Path
+import h5py
+from series import RunSeries
+import matplotlib.pyplot as plt
+import numpy as np
+from paths import rei640_reo_1359
+
+plt.style.use('prl')
+path = RunSeries(rei640_reo_1359, 'rei640_reo_-1359')
+
+filetype = Path("slices/slices_s1.h5")
+path.data = []
+path.r = []
+
+for i, r in enumerate(path.run_list):
+    filename = path.root_dir/Path(r)/filetype
+    with h5py.File(filename, "r") as df:
+        path.data.append(df['tasks/v_tot'][:,0,0,:])
+        path.r.append(df['scales/r/1.0'][:])
+    label = path.Lambdaz[i]
+
+    if label is None:
+        label = '\infty'
+    print(f"Lambda = {label} number of samples = {path.data[i].shape[0]}")
+    plt.plot(path.r[i],path.data[i][-25:,:].mean(axis=0), label=f"$\Lambda = {label}$")
+plt.xlabel(r"$r/d$")
+plt.ylabel(r"$\left< v \right>$")
+plt.xlim(path.r[0].min(),path.r[0].max())
+plt.legend()
+plt.tight_layout()
+plt.savefig(f"../figs/{path.name}_vmean_profile.pdf")

### 2022-07-30 python/series.py

@@ -7,8 +7,9 @@ class RunSeries:
     """Tool to manage a series of runs.
 
     """
-    def __init__(self, run_list, root_dir="results/"):
+    def __init__(self, run_list, name=None, root_dir="results/"):
         self.run_list = run_list
+        self.name = name
         self.root_dir = Path(root_dir)
         self.sim_time = []
         self.data = {}
@@ -19,6 +20,18 @@ class RunSeries:
         self.Re = []
         self.Re_o = []
         self.Lz = []
+        self.Lambdaz = []
+        self.Lambdat = []
+
+        for r in self.run_list:
+            eta, Re, Lz, mu, R1, R2, Omega1, period, Re_o, Lambdaz, Lambdat = self.parse_run_name(r)
+            self.eta.append(eta)
+            self.Lz.append(Lz)
+            self.mu.append(mu)
+            self.Re.append(Re)
+            self.Re_o.append(Re_o)
+            self.Lambdaz.append(Lambdaz)
+            self.Lambdat.append(Lambdat)
 
     def parse_run_name(self, fn):
         
@@ -26,6 +39,17 @@ class RunSeries:
         Re = float(re.search("re_([\d.e+-]+)",fn).group(1))
         Lz = float(re.search("Gamma_([\d.e+-]+)",fn).group(1))
         mu = float(re.search("mu_([\d.e+-]+)",fn).group(1))
+        Lambdaz = re.search("Lambdaz\_(\d+)",fn)
+        if Lambdaz:
+            Lambdaz = int(Lambdaz.group(1))
+        else:
+            Lambdaz = None
+        Lambdat = re.search("Lambdat\_(\d+)",fn)
+        if Lambdat:
+            Lambdat = int(Lambdat.group(1))
+        else:
+            Lambdat = None
+
         # derived parameters
         R1 = eta/(1. - eta)
         R2 = 1./(1-eta)
@@ -33,18 +57,12 @@ class RunSeries:
         period = 2*np.pi/Omega1
         Re_o = Re*mu/eta
 
-        return eta, Re, Lz, mu, R1, R2, Omega1, period, Re_o
+        return eta, Re, Lz, mu, R1, R2, Omega1, period, Re_o, Lambdaz, Lambdat
         
     def load(self, filetype, field, concatenate=True):
         filetype = Path(filetype)
         self.data[field] = []
         for r in self.run_list:
-            eta, Re, Lz, mu, R1, R2, Omega1, period, Re_o = self.parse_run_name(r)
[…]

### 2022-07-30 paper/paper.tex
Updates from Overleaf

@@
 In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that contribute to the mean flow.
-Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes.
-GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows \cite{2016PhRvL.116u4501M,2017JFM...810..412T, 2022arXiv220505513M}.
+Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes (see \cite{2022arXiv220505513M} for a recent review).
+GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.
 
@@
 \subsection{Spiral Turbulence}
-Meseguer et al (2009 PRE): 
-\begin{itemize}
-    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
-    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
-\end{itemize}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{../figs/reo_rei_lsb.pdf}
+    \caption{A summary of spiral turbulence runs in the $\Reyn_o-\Reyn_i$ plane showing linear stability boundary.
+Solid circles shows points where we have conducted GQL analysis; crosses represent DNS steps along the path to subcritical behaviour.
+}
+    \label{fig:LSB}
+\end{figure}
+%Meseguer et al (2009 PRE): 
+%\begin{itemize}
+%    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
+%    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
+%\end{itemize}
 
-The spiral turbulence regime with $\mu < 0$ provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_1$ and $\Reyn_2$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.
+The spiral turbulence regime provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.
+Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
+From these two seed runs, we decrease $\Reyn_i$ in steps shown in figure~\ref{fig:LSB} where each point represents a DNS or GQL simulation run for one viscous time.
+The solid circles in figure~\ref{fig:LSB} represent runs where we performed GQL analyses.
 
@@
 The second set of runs will
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
+    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
+    \label{fig:urms_tz_rei700}
+\end{figure}
 
@@
 F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
+G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
 \end{tabular}
@@
 
-\funding{Who paid Morgan? Who paid me?}
+\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. }
 

### 2022-07-31 paper/paper.tex
Updates from Overleaf

@@
 We solve the incompressible Navier-Stokes equations with unit density for $\mathbf{u}' = \mathbf{u} - \mathbf{U}$,
-where $\mathbf{U} = A r + B/r \hat{\mathbf{e}}_\phi$ is circular Couette flow (CCF here after),
+where $\mathbf{U} = A r + B/r \hat{\mathbf{e}}_\phi$ is circular Couette flow (CCF),
 \begin{align}
-      \frac{\partial \mathbf{u}}{\partial t} + \mathbf{u} \cdot \nabla \mathbf{u} &= -\nabla p + \nu \nabla^2\mathbf{u}\\
-      \nabla \cdot \mathbf{u} &= 0.
+\label{eqn:NS}
+      \frac{\partial \mathbf{u'}}{\partial t} + \mathbf{U} \cdot \nabla \mathbf{u'} + \mathbf{u'} \cdot \nabla \mathbf{U} + \mathbf{u'} \cdot \nabla \mathbf{u'} &= -\nabla p + \nu \nabla^2\mathbf{u'}\\
+      \nabla \cdot \mathbf{u'} &= 0.
 \end{align}
+In order to implement the GQL approximation, we introduce a projection operator to divide $\mathbf{u'}$ into two subsets,
+\begin{align}
+    \mathbf{u'}_l = \left< \mathbf{u'}\right>_l &= \sum_{m = -\Lambda_\theta}^{\Lambda_\theta} \sum_{k = -\Lambda_z}^{\Lambda_z} \mathbf{\hat{u}'}_{mk}(r) e^{i (m \theta + 2\pi k z/L_z)}\\
+    \mathbf{u'}_h = \mathbf{u'} - \mathbf{u'}_l.
+\end{align}
+GQL can then be written as 
+\begin{align}
+\label{eqn:GQL}
+      \frac{\partial \mathbf{u'}}{\partial t} + \mathbf{U} \cdot \nabla \mathbf{u'} + \mathbf{u'} \cdot \nabla \mathbf{U} +\nabla p - \nu \nabla^2\mathbf{u'} = &- \left<\mathbf{u'}_l \cdot \nabla \mathbf{u'}_l +\mathbf{u'}_h \cdot \nabla \mathbf{u'}_h\right>_l\\
+      & +\left<\mathbf{u'}_l \cdot \nabla \mathbf{u'}_h + \mathbf{u'}_h \cdot \nabla \mathbf{u'}_l \right>_h\\
+      \nabla \cdot \mathbf{u'} = & 0,
+\end{align}
+where we have written all linear terms on the left hand side.
+This form is equivalent to those given in prior works (for example \cite{2017JFM...810..412T}), where separate equations are written for the low and high modes, respectively.
 
-We use the Dedalus framework \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
-
-We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_1 = R_1 \Omega_1$.
-Taylor-Couette flow can be parameterised in terms of four parameters, traditionally the inner cylinder Reynolds number $\Reyn_1$, the ratio of inner to outer radii $\eta \equiv R_1/R_2$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_2/\Omega_1$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$.
+We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_i = R_i \Omega_i$.
+Taylor-Couette flow can be parameterised in terms of four parameters, the inner cylinder Reynolds number $\Reyn_i$, the ratio of inner to outer radii $\eta \equiv R_i/R_o$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_o/\Omega_i$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$.
 In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length.
-We will also refer to the two Reynolds numbers $\Reyn_1$, $\Reyn_2$, related by
+We will also refer to the two Reynolds numbers $\Reyn_i$, $\Reyn_2o$, related by
 \begin{equation}
@@
 \end{equation}
-
-It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_1$.F
+It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
+We use the Dedalus framework \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
 \section{Linear Stability}
@@
 \end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
+    \includegraphics[width=0.5\textwidth]{}
+    \caption{Perturbation kinetic energy as a function of time for $\Reyn_i = 900, \Reyn_o=-3398$.
+Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$, we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$.
+Note that all of the reduced models begin high amplitude oscillations.}
+    \label{fig:ke_vs_t_rei900}
+\end{figure}
 %Meseguer et al (2009 PRE): 
@@
 \label{sec:nonlinear}
-Taylor-Couette flow is characterized by $\Reyn_i$, $\Reyn_o$, $\eta$, $\Gamma$. For all runs in this paper, the outer cylinder is stationary, so $\Reyn_o = 0$.
 

### 2022-07-31 paper/paper.tex

@@
 \author{%%%% Author details
-Jeffrey S. Oishi$^{1}$ and Morgan Baxter$^{1,2}$}
+Jeffrey S. Oishi and Morgan Baxter}
 
 %%%%%%%%% Insert author address here
-\address{$^{1}$Department of Physics and Astronomy, Bates College, Lewiston, ME USA\\
-$^{2}$Company Name, Whereever, USA}
+\address{Department of Physics and Astronomy, Bates College, Lewiston, ME USA}
 
@@
 It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
-We use the Dedalus framework \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations. Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
+We use the Dedalus framework (version 2) \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations.
+Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
 \section{Linear Stability}
@@
 
-\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
-Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
-\begin{itemize}
-    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
-    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
-    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
-    \item Compare GQL performance at various cutoffs
-    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
-\end{itemize}
+%\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
+%Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
+%\begin{itemize}
+%    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
+%    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
+%    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
+%    \item Compare GQL performance at various cutoffs
+%    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
+%\end{itemize}
 \subsection{Spiral Turbulence}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
-    \includegraphics[width=0.5\textwidth]{}
-    \caption{Perturbation kinetic energy as a function of time for $\Reyn_i = 900, \Reyn_o=-3398$.
-Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$, we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$.
-Note that all of the reduced models begin high amplitude oscillations.}
-    \label{fig:ke_vs_t_rei900}
-\end{figure}
 %Meseguer et al (2009 PRE): 
@@
 
-The spiral turbulence regime provides a unique opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical routes to spiral turbulence can be selected.
+The spiral turbulence regime provides an ideal opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical manifestations of spiral turbulence can be selected.
 Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
@@
 
-\section{Runs}
+\section{Results}
 \label{sec:nonlinear}
-
-The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}, particularly section 4 of \cite{1984JFM...146...65M}.
+%The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}, particularly section 4 of \cite{1984JFM...146...65M}.
 
@@
     \label{fig:urms_tz_rei700}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
+    \end{minipage}
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
+    \end{minipage}
+    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$.
+On the left,  $\Reyn_i = 900$ is above the linear stability boundary, while the right shows $\Reyn_i = 700$, below the stability boundary.
+Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$.
+For $\Reyn_i = 700$, we also use $\Lambda = 20, 30$ At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations.
+}
+    \label{fig:ke_vs_t_rei900}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figs}
+    \caption{Caption}
+    \label{fig:my_label}
 \end{figure}
@@
 
-\section{Resolution}
-Might compare poorly resolved DNS to low order GQL as done in\cite{2017JFM...810..412T}. Note that in poorly resolved TC DNS, one gets spurious spiral modes.
+%\section{Resolution}
+%Might compare poorly resolved DNS to low order GQL as done in\cite{2017JFM...810..412T}. Note that in poorly resolved TC DNS, one gets spurious spiral modes.
 
@@
 
-\appendix{Odds and ends}
-\begin{itemize}
-    \item On the $\Reyn_c(m=1)/\Reyn_c(m=0) \simeq 1.008$ issue...there's a discussion in  Krueger, Gross, Di Prima (1966), p 523-524, but it references a solution in a conference proceeding that is difficult to find.
+%\appendix{Odds and ends}
+%\begin{itemize}
+%    \item On the $\Reyn_c(m=1)/\Reyn_c(m=0) %\simeq 1.008$ issue...there's a discussion in  Krueger, Gross, Di Prima (1966), p 523-524, but it references a solution in a conference proceeding that%is difficult to find.
 I have some memory of it being discussed in a more recent paper, but I can't find it anywhere...
-\end{itemize}
-\enlargethispage{20pt}
+%\end{itemize}
+%\enlargethispage{20pt}
 

### 2022-07-31 python/plot_rei900_snapshots_figure.py
added rei90 snapshot figure and code.

@@ -1,0 +1,74 @@
+from pathlib import Path
+import h5py
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import AxesGrid
+import numpy as np
+import paths
+from series import RunSeries
+from dedalus.extras import plot_tools
+plt.style.use('prl')
+
+#path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
+path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
+
+aspect=0.9 #np.sqrt(2)
+w = 5
+h = aspect*w
+
+def load_data(df, Rmid, index=-1):
+    dset = np.sqrt(df['tasks']['u_r_slice'][index, :,:,0]**2 + df['tasks']['v_r_slice'][index, :,:,0]**2 + df['tasks']['w_r_slice'][index, :,:,0]**2)
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),0]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/u_r_slice'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+
+
+fig = plt.figure(figsize=(w, h))
+grid = AxesGrid(fig, 111,
+                nrows_ncols=(3, 2),
+                axes_pad=0.05,
+                cbar_mode='single',
+                cbar_location='top',
+                cbar_pad=0.
+                )
+
+
+for i,ax in enumerate(grid):
+    if i != 4:
+        ax.set_axis_off()
+    else:
+        ax.set_xlabel(r"$R_{mid}\theta$", fontsize=14)
+        ax.set_ylabel("z", fontsize=14)
+        ax.tick_params(axis='both', which='major', labelsize=14)
+    r = path.run_list[i]
+    eta = path.eta[i]
+    print(f"plotting {r}")
+
+    filetype = 'slices/slices_s2.h5'
+    if not (path.root_dir/Path(r)/filetype).exists():
+        filetype = 'slices/slices_s1.h5'
+    filename = path.root_dir/Path(r)/filetype
+
+    Rmid = eta/(1-eta) + 0.5
+    with h5py.File(filename, "r") as df:
+        xgrid, ygrid, data = load_data(df, Rmid)
+
[…]

### 2022-08-01 python/plot_rei900_story_figure.py
added story figure for rei=900 Lambda=5

@@ -1,0 +1,98 @@
+from pathlib import Path
+import h5py
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import AxesGrid
+import numpy as np
+import paths
+from series import RunSeries
+from dedalus.extras import plot_tools
+plt.style.use('prl')
+
+#path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
+path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
+
+
+aspect=0.7 #np.sqrt(2)
+w = 10
+h = aspect*w
+
+def load_data(df, Rmid, index=-1):
+    dset = np.sqrt(df['tasks']['u_r_slice'][index, :,:,0]**2 + df['tasks']['v_r_slice'][index, :,:,0]**2 + df['tasks']['w_r_slice'][index, :,:,0]**2)
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),0]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/u_r_slice'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+
+rect = [0.5,0.15,0.4,0.8]
+fig = plt.figure(figsize=(w, h))
+ekin_ax = fig.add_axes([0.1,0.15,0.4,0.8])
+grid = AxesGrid(fig, rect,
+                nrows_ncols=(3, 1),
+                axes_pad=0.05,
+                # cbar_mode='single',
+                # cbar_location='top',
+                # cbar_pad=0.
+                )
+filetype = Path("scalar/scalar.h5")
+path.data = []
+path.t = []
+t_visc = path.Re[0]
+print(f"t_visc = {t_visc}")
+path_i = 4
+indices = [81,89,-4]
+
+r = path.run_list[path_i]
+eta = path.eta[path_i]
+label = path.Lambdaz[path_i]
+filename = path.root_dir/Path(r)/filetype
+with h5py.File(filename, "r") as df:
+    path.data.append(df['tasks/pertubation_KE'][:,0,0,0])
+    path.t.append(df['scales/sim_time'][:])
+
+
+if label is None:
+    label = '\infty'
+
+ekin_ax.plot(path.t[0]/t_visc, path.data[0], label=f"$\Lambda = {label}$")
+ekin_ax.set_xlabel(r"$t/\tau_\nu$")
[…]

### 2022-08-01 paper/paper.tex
Updates from Overleaf

@@
 \begin{abstract}
-
+Taylor-Couette flow is well known to admit a spiral turbulence state in which laminar and turbulent patches coexist around the cylinder.
+This flow state is quite complex, with delicate internal structure, and it can be traced into certain regimes of linear stability.
+This behavior is believed to be connected to the non-normality of the linear operator, which is itself a function of the control parameters.
+Using spiral turbulence in both linearly stable and unstable regimes, we investigate the effectiveness of the generalized quasilinear approximation (GQL), an extension of quasilinear theory designed to capture the essential aspects of turbulent flows.
+We find that GQL performs much better in the supercritical regime than the subcritical.
+By including only a small number of modes into the non-linear interactions, GQL simulations maintain a turbulent-like state when in the supercritical regime.
+However, a much larger number is required to avoid returning to the laminar state when in the subcritical regime.
 \end{abstract}
@@
 
-\begin{fmtext}
-\end{fmtext}
+%\begin{fmtext}
+%\end{fmtext}
 %%%%%%%%%%%%%%% End of first page %%%%%%%%%%%%%%%%%%%%%
@@
 Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$.
-This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the typical cascade found at high $\Reyn$.
+This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$.
 %GQL has been show to outperform QL models in numerous highly turbulent flows, driven by $\beta$-effects on rotating spheres, rotating convection, rotating plane Couette flow (pCf, here after). 
 One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
-In the linearly stable regions, transient growth has been observed in TC flow.
-As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow can be tuned to control the degree of non-normality, that is its commutator with its adjoint
+In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}.
+We can always write the Navier-Stokes equations in schematic form as
 \begin{equation}
-    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn, \Gamma)
+    \label{eqn:NS-schematic}
+    \frac{\partial \mathbf{q}}{\partial t} + \Lop \mathbf{q} = \mathcal{N}(\mathbf{q}),
+\end{equation}
+where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u}  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
+As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow is a function of its control parameters.
+That is, its commutator with its adjoint
+\begin{equation}
+    [\Lop^\dagger, \Lop] = f(\mu, \eta, \Reyn_i, \Gamma)
 \end{equation}
 can be zero or non-zero. 
-This provides a very useful arena to test the validity of GQL in a system that admits both normal and non-normal operators.
+This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.
 
-We pursue the flow into the \emph{spiral turbulence} flow regime, and consider the effectiveness of GQL in each case.
-The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a subcritical transition to spiral turbulence when the cylinders were counter-rotating.
+As a first experiment to this end, we consider the \emph{spiral turbulence} flow regime.
+The classic paper by Coles \cite{1965JFM....21..385C} was the first to note a hysteretic transition to spiral turbulence when the cylinders were counter-rotating: an experiment at fixed $\Reyn_o$ will have spiral turbulence at a certain $\Reyn_i$ if $\Reyn_i$ is reached by \emph{decreasing} from a higher value, but will not have spiral turbulence at the same $\Reyn_i$ if it is reached by \emph{increasing} from a lower value.
 More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}.
-Using these prior works as a guide, we consider three cases
-\begin{enumerate}
-    \item Inner cylinder rotating only $\Reyn_o = 0$
-    \item Counter rotating cylinders with moderate outer cylinder rotation $\Reyn_o = -1339$
-    \item Counter rotating cylinders with high outer cylinder rotation $\Reyn_o = -3389$
-\end{enumerate}
-Case 1 has a purely normal linear operator, while the latter two feature moderate and significant non-normality respectively.
+In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarizes below the stability boundary when $\Reyn_o = -1200$.
+Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region.
+We then run a suite of GQL simulations in both the super- and subcritical regions.
+%Using these prior works as a guide, we consider three cases
+%\begin{enumerate}
+%    \item Inner cylinder rotating only $\Reyn_o = 0$
+%    \item Counter rotating cylinders with moderate outer cylinder rotation $\Reyn_o = -1339$
+%    \item Counter rotating cylinders with high outer cylinder rotation $\Reyn_o = -3389$
+%\end{enumerate}
+%Case 1 has a purely normal linear operator, while the latter two feature moderate and significant non-normality respectively.
 
@@
 This form is equivalent to those given in prior works (for example \cite{2017JFM...810..412T}), where separate equations are written for the low and high modes, respectively.
+For all runs in this paper, $\Lambda = \Lambda_z = \Lambda_\theta$, though we note it would be interesting to consider models quasilinear in $\theta$ (i.e.
+$\Lambda_\theta = 0$) with varying $\Lambda_z$.
+These correspond to the \emph{restricted non-linear} models used with success in shear flow turbulence \cite{2014PhFl...26j5112T, 2015PhFl...27j5104T, 2015PhFl...27a1702B}.
 
@@
 In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length.
-We will also refer to the two Reynolds numbers $\Reyn_i$, $\Reyn_2o$, related by
+We will also refer to the two Reynolds numbers $\Reyn_i$, $\Reyn_o$, related by
 \begin{equation}
-    \frac{\Reyn_2}{\Reyn_1} = \frac{\mu}{\eta}.
+    \frac{\Reyn_o}{\Reyn_i} = \frac{\mu}{\eta}.
 \end{equation}
 It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
-We use the Dedalus framework (version 2) \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations.
+
+We use the pseudo-spectral Dedalus framework (version 2) \cite{2020PhRvR...2b3068B} for all linear and non-linear calculations.
+We discretize the three velocity components and pressure onto a basis of Chebyshev polynomials in the radial direction, while $\theta$ and $z$ use Fourier bases.
+To evolve the solution forward in time, we use the implicit-explicit Runga-Kutta RK443 time stepper; linear terms are implicitly timestepped while non-linear terms are explicit.
+Our initial value solver has been extensively tested for TC flow against the linear and non-linear test cases in \cite{1984JFM...146...45M, 1984JFM...146...65M} including linear growth rates, non-linear wave speeds, and torque measurements; our spiral turbulence DNS results also agree with those in \cite{2009PhRvE..80d6315M,2009PhRvE..80f7301D,2011JFM...668..150D}.
+Resolutions and geometric parameters for all runs are given in table~\ref{tab:simulations}.
+We have confirmed our simulations are resolved by ensuring that the highest wavenumber components have spectral power at least $10^5$ smaller than the most energetic scale.
+
 Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
@@
 \label{sec:linear}
-\subsection{Non-Axisymmetric Instability with $\mu = 0$}
-%HAHAH This is all laid out in Krueger, Gross, Di Prima (1966), p 523-524
-%The axisymmetric test case of \cite{1984JFM...146...65M} has $\mu = 0$, $\eta = 0.875$, $\Reyn = 139.31064$, and %$k_{min} = 2\pi/L_z = 2.513$ ($L_z = 2.5$).
-Interestingly, table~\ref{tab:critical_m} shows that circular Couette %flow is linearly unstable to all $m \leq 4$ modes at this $\Reyn$.
-%
-%A natural assumption is that one would not expect to see axisymmetric flow at these parameters.
-Fully non-linear %simulations initialized from very low-amplitude random, $\mathbf{\nabla \cdot u} = 0$ perturbations bear this out.
-%Yet, \cite{1984JFM...146...65M} makes clear reference to TVF being stable to perturbations at this flow.
-He also %claims that initial conditions including ``round-off noise of the computer'' lead to axisymmetric flow.
-This is %quite unlikely to be true.
-%
-%% the two runs (axisymmetric TVF and perturbed, restarted run) are on Leavitt
-%One is naturally led to the question of how axisymmetric TVF is seen in experiment if $\Reyn_{c,m=1}/\Reyn_{c, %m=0} \simeq 1.008$: $m=1$ modes are unstable at a supercriticality of less than 1\%! 
-%The answer becomes obvious if one considers the usual experimental procedure: spin up a fluid from rest to a %$\Reyn < \Reyn_{c,m=0}$, allow Couette flow to be established, then spin up until the first bifurcation occurs.
-%Allow the TVF to settle and proceed upwards.
-This will indeed produce an axisymmetric TVF flow that is %\emph{itself stable against $m\ne 0$ perturbations}!
-At $\Reyn = 138.31$, axisymmetric TVF flow can be perturbed %with gaussian noise and will remain axisymmetric.
-%
-%\begin{table}[!h]
-%\caption{Critical inner $\Reyn_i$ and $k_z$ as a function of $m$ for $\eta = 0.875$, $\mu = 0$}%%%Table caption %goes here
-%\label{tab:critical_m}
-%\begin{tabular}{rll}%%%The number of columns has to be defined here
-%\hline
-%m & $\Reyn_{ic}$ & $k_z$ \\
-%\hline
-%0 & 118.1 & 3.12947\\
-%1 & 119.0 & 3.14374\\
-%2 & 121.5 & 3.18888\\
-%3 & 126.17 & 3.25852\\
-%4 & 133.32 & 3.35672\\
-%5 & 144.20 & 3.47885\\
-%\end{tabular}
-%\vspace*{-4pt}
-%\end{table}%%%End of the table
-
-The axisymmetric test case of \cite{1984JFM...146...65M} has $\mu = 0$, $\eta = 0.875$, $\Reyn = 139.31064$, and $k_{min} = 2\pi/L_z = 2.513$ ($L_z = 2.5$).
-Interestingly, circular Couette flow is linearly unstable to all $m \leq 4$ modes at this $\Reyn$.
-In fact, linear stability analysis shows that $\Reyn_{c,m=1}/\Reyn_{c, m=0} \simeq 1.008$: $m=1$ modes are unstable at a supercriticality of less than 1\%. 
-One is naturally led to the question of how axisymmetric TVF is seen in experiment!
-The answer becomes clear if one considers the usual experimental procedure: spin up a fluid from rest to a $\Reyn < \Reyn_{c,m=0}$, allow Couette flow to be established, then spin up until the first bifurcation occurs.
-Allow the TVF to settle and proceed upwards.
-This will indeed produce an axisymmetric TVF flow that is \emph{itself stable against $m\ne 0$ perturbations}!
-At $\Reyn = 138.31$, axisymmetric TVF flow can be perturbed with gaussian noise and will remain axisymmetric.
-
-This is important for understanding the bifurcations from TVF to WVF: starting from $\Reyn < \Reyn_{W}$ with random initial conditions, one will find a non-axisymmetric flow.
-
-%\subsection{$\Reyn_2 = -1000$, $\eta = 0.883$: Non-modal growth and GQL}
-%Next, we look at the case of $\Reyn_2 = -1000$. Setting $\eta = 0.883$ as in Andereck et al (1986), we consider the performance of GQL in both the linearly stable and unstable regime
-%\begin{itemize}
-%    \item look at pseudospectra for $\Reyn_1 = 450$, $\Reyn_1 = 400$ (below and above critical)
-%    \item compare pseudospectra between $\Reyn_2 = 0$ and $\Reyn_2 = -1000$
-%    \item do same thing at smaller $\Reyn_2$ to compare with \cite{2002PhFl...14.3475H}
-%    \item Compare GQL performance at various cutoffs
-%    \item Identify non-normality parameter for TC $\Lop$ for both linearly unstable state and non-linearly modified background
-%\end{itemize}
-\subsection{Spiral Turbulence}
+In order to determine the linear stability bound, we find the minimum $\Reyn_i$ for a given $\Reyn_o$ by solving the eigenvalue problem for equations~\ref{eqn:NS}.
+We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, most unstable mode is never axisymmetric.
+At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset, and
 \begin{figure}
@@
 \end{figure}
-
-%Meseguer et al (2009 PRE): 
-%\begin{itemize}
-%    \item $\Reyn_o = -1200, \Reyn_i = 600$, $\eta=0.883$, $\Gamma = 29.9$
-%    \item $(N_r, N_\theta, N_z) = (20,220,220)$ HAHAHA. No way...more like $(64, 512, 512)$
-%\end{itemize}
 
@@
 
-First, consider a set of runs with $\mu = 0$.
+%First, consider a set of runs with $\mu = 0$.
 This is a standard, linearly unstable case.
@@
 The second set of runs will
-
+We first consider the supercritical case with $\Reyn_i = 900$.
+We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
+Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$.
+The first important thing to note is that none of these low-order models are able to maintain the spiral structure.
+Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs.
+The supercritical case shows significant oscillations at all $\Lambda$.
+Focusing on $\Lambda=1$, we see that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns: the interpenetrating spirals (IPS) and intermittency regimes (INT) originally identified in experiments \cite{1984JFM...146...45M} and subsequently found in simulations (e.g.
+\cite{2009PhRvE..80d6315M}).
 \begin{figure}
@@
     \label{fig:urms_tz_rei700}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figs/rei900_snapshots.pdf}
+    \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$.
+The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$.
+Even at $\Lambda = 10$, spiral turbulence is not apparent.}
+    \label{fig:rei900_snapshots}
 \end{figure}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_vmean_profile.pdf}
+    \caption{Mean deviation from CCF for $\Reyn_i = 900$, $\Reyn_o=-3398$. The heavy line is DNS, the orange line is QL, and all others are GQL with $\Lambda = 1,3,5,10$.}
+    \label{fig:rei900_vmean}
+\end{figure}
+
 \begin{table}[!h]
@@
 \hline
-A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
-B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
-C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
-\hline
-D & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
-E & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Subcritical SPT\\
-\hline
-F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
-G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
+%A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
+%B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
+%C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
+%\hline
+A & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
+B & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
+%\hline
+%F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
+%G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
 \end{tabular}

### 2022-08-01 python/TC_pseudospectrum.py
added pseudospectra figures.

@@ -35,7 +35,7 @@ nr=int(args['--nr'])
 m = int(args['--m'])
 alpha = float(args['--kz'])
 
-filename = "../figs/TC_pseudospectrum_Re1_{:5.2f}_eta_{:6.4f}_mu_{:6.4f}_nr_{:d}_m_{:d}_kz_{:5.2f}.png".format(Re1, eta, mu, nr, m, alpha)
+filename = "../figs/TC_pseudospectrum_Re1_{:5.2f}_eta_{:6.4f}_mu_{:6.4f}_nr_{:d}_m_{:d}_kz_{:5.2f}.pdf".format(Re1, eta, mu, nr, m, alpha)
 
 def energy_norm(Q1, Q2):
     u1 = Q1['u']
@@ -140,17 +140,21 @@ k = 25
 
 psize = 100
 ps_real = np.linspace(-4,2,psize)
-ps_imag = np.linspace(-3,3,psize)
+ps_imag = np.linspace(-4,4,psize)
 ep.calc_ps(k, (ps_real, ps_imag), inner_product=energy_norm,maxiter=20)
 
 clevels = np.linspace(-1.8,0,7)
 print("contour levels: ", clevels)
 plt.figure(figsize=(5,5))
 plt.scatter(ep.evalues_good.imag, ep.evalues_good.real)
-plt.xlim(-1.5,1.5)
-plt.ylim(-2,1)
+print(f"number of good evals: {len(ep.evalues_good)}")
+#plt.xlim(-1.5,1.5)
+#plt.xlim(-1.5,5)
+#plt.ylim(-10,1)
 plt.axhline(0,color='k',alpha=0.3)
-plt.contour(ep.ps_imag, ep.ps_real, np.log10(ep.pseudospectrum.T),levels=clevels,colors='k')
+CS2 = plt.contour(ep.ps_imag, ep.ps_real, np.log10(ep.pseudospectrum.T),levels=clevels,colors='k')
+ax2 = plt.gca()
+ax2.clabel(CS2, CS2.levels)
 plt.xlabel(r"$\omega$")
 plt.ylabel(r"$\gamma$")
 plt.tight_layout()

### 2022-08-01 paper/paper.tex

@@
 This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$.
-%GQL has been show to outperform QL models in numerous highly turbulent flows, driven by $\beta$-effects on rotating spheres, rotating convection, rotating plane Couette flow (pCf, here after). 
 One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
@@
 We then run a suite of GQL simulations in both the super- and subcritical regions.
-%Using these prior works as a guide, we consider three cases
-%\begin{enumerate}
-%    \item Inner cylinder rotating only $\Reyn_o = 0$
-%    \item Counter rotating cylinders with moderate outer cylinder rotation $\Reyn_o = -1339$
-%    \item Counter rotating cylinders with high outer cylinder rotation $\Reyn_o = -3389$
-%\end{enumerate}
-%Case 1 has a purely normal linear operator, while the latter two feature moderate and significant non-normality respectively.
 
@@
 Resolutions and geometric parameters for all runs are given in table~\ref{tab:simulations}.
+\begin{table}[!h]
+\caption{Simulations}
+\label{tab:simulations}
+\begin{tabular}{rlllllllll}
+\hline
+$\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
+\hline
+%A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
+%B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
+%C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
+%\hline
+900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
+700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
+%\hline
+%F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
+%G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
+\end{tabular}
+\vspace*{-4pt}
+\end{table}%%%End of the table
 We have confirmed our simulations are resolved by ensuring that the highest wavenumber components have spectral power at least $10^5$ smaller than the most energetic scale.
@@
 Pseudospectra and critical parameters are calculated using eigentools \cite{2021JOSS....6.3079O}.
+
 \section{Linear Stability}
@@
 \label{sec:nonlinear}
-%The ($\eta$, $\Gamma$) choice is motivated by \cite{1984JFM...146...45M,1984JFM...146...65M}, particularly section 4 of \cite{1984JFM...146...65M}.
-
-%First, consider a set of runs with $\mu = 0$.
-This is a standard, linearly unstable case.
-Here, we choose three $\Reyn$ values, one in the WVF flow regime $\Reyn =750$, one in the MWF $\Reyn = 1500$, and one in the TTV regime $\Reyn = 2000$.
-We perform two sets of runs, each varying $\Lambda = {0, 1, 2, 6, 10, 20}$ for both the $z$ and $\theta$ directions.
-The first set of runs is a cold start from identical initial conditions.
-This allows us to understand the ability of GQL to find certain flow states.
-The second set of runs will
-We first consider the supercritical case with $\Reyn_i = 900$.
-We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
-Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$.
-The first important thing to note is that none of these low-order models are able to maintain the spiral structure.
-Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs.
-The supercritical case shows significant oscillations at all $\Lambda$.
-Focusing on $\Lambda=1$, we see that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns: the interpenetrating spirals (IPS) and intermittency regimes (INT) originally identified in experiments \cite{1984JFM...146...45M} and subsequently found in simulations (e.g.
-\cite{2009PhRvE..80d6315M}).
+Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the rms velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$.
+Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$.
 \begin{figure}
@@
 \end{figure}
-
+\subsection{Supercritical Spiral Turbulence}
+We first consider the supercritical case with $\Reyn_i = 900$.
+We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
+Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$.
+The first important thing to note is that none of these low-order models are able to maintain the spiral structure.
+However, even including a single additional low mode, $\Lambda = 1$, shows a significant improvement in flow morphology compared to QL.
+It is quite interesting to note that QL \emph{does} feature some degree of spatial intermittency--there are clearly separated regions hosting spiral waves interspersed with featureless, laminar patches.
 \begin{figure}
@@
 
+Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
+The supercritical case shows significant oscillations at all $\Lambda$. 
 \begin{figure}
@@
 \end{figure}
+Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g.
+\cite{2009PhRvE..80d6315M}).
+At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating.
+As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT).
+Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches is closests approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.
 
+This is notable for a few reasons.
+At $\Reyn_i=900, \Reyn_o=-3398$ the system is not in a hysteretic region of parameter space.
+Thus, we do not expect this behaviour to be due to GQL picking up other, coexisting solutions.
+It also highlights the delicate balance an intermittent laminar/turbulent solution represents.
+The turbulence in this solution has a very broad range of scales, as evidenced by figure~\ref{fig:urms_tz_rei700}.
+Understanding how it exchanges energy with the mean flow is crucial to understanding both the saturation of the underlying linear instability as well as the maintenance of the non-linear state.
 \begin{figure}
     \centering
-    \includegraphics{figs}
-    \caption{Caption}
-    \label{fig:my_label}
+    \includegraphics[width=0.9\textwidth]{figs/rei900_story.pdf}
+    \caption{Kinetic energy (left) and flow morphologies (right) for $\Reyn_i = 900$, $\Lambda = 5$ covering one full oscillation cycle.
+The panels on the right show three times corresponding to the vertical lines in the kinetic energy plot on the left.
+At the kinetic energy minimum (point 1), the flow shows a IPS-like morphology, while an INT burst occurs at the knee (point 2), and a more space-filling turbulent state occurs at maximum energy (point 3).}
+    \label{fig:rei900_lambda5_story}
 \end{figure}
@@
 
-\begin{table}[!h]
-\caption{Simulations}
-\label{tab:simulations}
-\begin{tabular}{rlllllllll}
-\hline
-Run & $\Reyn_i$ & $\Reyn_o$ & $\eta$ & $\Gamma$  & $N_r$ & $N_\theta$ & $N_z$& $\Lambda_\theta$, $\Lambda_z$ & Notes\\
-\hline
-%A & 750  & 0 & 0.875  & 3 & 32 & 64 & 64 & 0, 1, 3, 5 & WVF\\
-%B & 1000 & 0 & 0.875 & 3 & 32 & 256 & 64 & 0, 1, 3, 5 & MWF\\
-%C & 2000 & 0 & 0.875 & 3 & 32 & 512 & 64 & 0, 1, 3, 5 & TTV\\
-%\hline
-A & 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
-B & 700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
-%\hline
-%F & 640 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5 & Supercritical SPT\\
-%G & 480 & -1359 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 5 & Supercritical IPS\\
-\end{tabular}
-\vspace*{-4pt}
-\end{table}%%%End of the table
-
-%\section{Resolution}
-%Might compare poorly resolved DNS to low order GQL as done in\cite{2017JFM...810..412T}. Note that in poorly resolved TC DNS, one gets spurious spiral modes.
-
-%\section{Conjugate Stable Modes}
-%Conjugate stable modes?
-Typically, one considers the effect of the inviscid stable mode rather than the viscously damped mode.
-However, in the case of TC flow, the background flow profile could not be maintained by the cylinder walls in the limit $\nu \to 0$.
-However, because circular Couette flow is independent of $\Reyn$, probably the limit is OK?
-Can think about this.
-Obviously Rayleigh criterion applies in inviscid case, and this can be easily checked for any angular momentum profile $r^2 \Omega(r)$, so there is some merit to thinking about ideal case even in TC flow.
-Need to think more about this/run it by Adrian.
-
-%The output for figure is:\vspace*{-7pt}
-
-%\begin{figure}[!h]
-%\centering\includegraphics[width=2.5in]{xxxxxx.eps}
-%%% where xxxxxx name represents "figurename.eps"
-%\caption{Insert figure caption here}
-%\label{fig_sim}
-%\end{figure}
+\subsection{Subcritical spiral turbulence}
+In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$.
+Each is initialised with the last timestep of the prior run and run for one viscous time.
+Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study.
+HThe right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$.
+Our very preliminary data for $\Lambda = 30$ show that it recovers from the decay and begins to climb in energy again, similar to the early behaviour from the supercritical runs.\footnote{\textbf{Note to reviewers}: this simulation is still running as of the 31 July due date for the initial manuscript.
+By the time refereeing begins, we will update this plot and section.}
 
@@
 \label{sec:conclusion}
+We have presented a series of simulations demonstrating the performance of GQL at various cutoffs for Taylor-Couette flow in the spiral turbulence regime.
+GQL is a significant improvement over QL models in representing mean quantities, including the deviation from circular Couette flow.
+However, even at the highest cutoffs studied here, it does not retain an unambiguous spiral form in either the sub- or supercritical regimes.
+Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}.
+This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realizations \cite{2022arXiv220505513M}.
 
-%\appendix{Odds and ends}
-%\begin{itemize}
-%    \item On the $\Reyn_c(m=1)/\Reyn_c(m=0) %\simeq 1.008$ issue...there's a discussion in  Krueger, Gross, Di Prima (1966), p 523-524, but it references a solution in a conference proceeding that%is difficult to find.
-I have some memory of it being discussed in a more recent paper, but I can't find it anywhere...
-%\end{itemize}
-%\enlargethispage{20pt}
+However, we have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow.
+In future work, we will consider the ability of GQL to maintain exact coherent states identified by Deguchi et al \cite{2014PhRvL.112r4502D} and the dynamics of which were studied recently by Wang and coworkers \cite{2022arXiv220712990W}.
+Understanding how these interactions work to shape subcritical TC turbulence represents a promising new avenue to understanding how ECS are maintained and how they in turn sustain turbulence below linear stability bounds.
+In particular, one can begin to understand the relative importance of
 
-%\ethics{Insert ethics statement here if applicable.}
-
-\dataccess{Insert details of how to access any supporting data here.}
+\dataccess{All code for non-linear initial value problems, linear eigenvalue and non-modal stability analyses, and plotting are available at \url{https://github.com/jsoishi/GQL_TC}.
+Dedalus version 2 is available at \url{https://github.com/DedalusProject/dedalus/tree/v2_master}.
+eigentools is available at \url{https://github.com/DedalusProject/eigentools}.
+Simulation data is available upon request.}
 
 \aucontribute{JSO wrote the manuscript, designed the study, wrote the GQL projection operators, and implemented the original Taylor-Couette Dedalus script.
-MB added the GQL terms to the TC script, developed data analysis techniques, and performed preliminary simulations.
-Both authors edited the manuscript.}
+MB added the GQL terms to the TC script, developed data analysis techniques, and performed preliminary simulations.}
 
@@
 
-\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. }
+\funding{JSO was supported by NASA LWS grant number NNX16AC92G and NASA HTMS grant number 80NSSC20K1280. MB was supported in part by a Bates College Travel Grant.}
 

### 2022-08-01 python/LSB.py
updated LSB

@@ -8,11 +8,11 @@ our_path=  np.loadtxt('data/our_path_1.csv',delimiter=',')
 our_path_2=np.loadtxt('data/our_path_2.csv',delimiter=',')
 
 # conversion factors for Coles (1965)
-# ri_c = 5.
-# ro_c = 11.44/2
-# d_c = ro_c - ri_c
-# ifactor = d_c/ri_c
-# ofactor = d_c/ro_c
+ri_c = 5.
+ro_c = 11.44/2
+d_c = ro_c - ri_c
+ifactor = d_c/ri_c
+ofactor = d_c/ro_c
 
 # coles_H1z = np.loadtxt("data/coles_1965_zoom.csv",delimiter=',')
 # cz_x = ofactor*(coles_H1z[:,2] + coles_H1z[:,0])/2
@@ -20,15 +20,15 @@ our_path_2=np.loadtxt('data/our_path_2.csv',delimiter=',')
 # cz_xerr = ofactor*(coles_H1z[:,2] - coles_H1z[:,0])/2
 # cz_yerr = ifactor*(coles_H1z[:,3] - coles_H1z[:,1])/2
 
-# andereck = np.loadtxt("data/andereck_H2.csv",delimiter=',')
-# andereck_x    = (andereck[:,2] + andereck[:,0])/2
-# andereck_y    = (andereck[:,3] + andereck[:,1])/2
-# andereck_xerr = (andereck[:,2] - andereck[:,0])/2
-# andereck_yerr = (andereck[:,3] - andereck[:,1])/2
+andereck = np.loadtxt("data/andereck_H2.csv",delimiter=',')
+andereck_x    = (andereck[:,2] + andereck[:,0])/2
+andereck_y    = (andereck[:,3] + andereck[:,1])/2
+andereck_xerr = (andereck[:,2] - andereck[:,0])/2
+andereck_yerr = (andereck[:,3] - andereck[:,1])/2
 
-# coles_H1 = np.loadtxt("data/coles_1965_H1.csv",delimiter=',')
-# c_xerr = ofactor*(coles_H1[:,2] - coles_H1[:,0])
-# c_yerr = ifactor*(coles_H1[:,3] - coles_H1[:,1])
+coles_H1 = np.loadtxt("data/coles_1965_H1.csv",delimiter=',')
+c_xerr = ofactor*(coles_H1[:,2] - coles_H1[:,0])
+c_yerr = ifactor*(coles_H1[:,3] - coles_H1[:,1])
 
[…]

### 2022-08-01 paper/paper.tex
Updates from Overleaf

@@
 \end{figure}
+The spiral turbulence regime provides an ideal opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical manifestations of spiral turbulence can be selected.
 
-The spiral turbulence regime provides an ideal opportunity to test GQL in an environment with both spatiotemporal patterns and a tunable bifurcation: by choosing $\Reyn_i$ and $\Reyn_o$ appropriately, both subcritical and supercritical manifestations of spiral turbulence can be selected.
-Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
+Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ 
+%for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and 
+start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
 From these two seed runs, we decrease $\Reyn_i$ in steps shown in figure~\ref{fig:LSB} where each point represents a DNS or GQL simulation run for one viscous time.
@@
 
+In order to demonstrate the effects non-normality plays, figure~\ref{fig:pseudospectra} shows both the modal spectrum (in blue points) and the $\epsilon-$pseudospectrum for the supercritical and subcritical $\Reyn_i$.
+The contour value $\epsilon^{-1}$ represents the set of values
+\begin{equation}
+    \lVert (\sigma \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
+\end{equation}
+for a given complex number $\sigma = \gamma + i\omega$ here representing the eigenvalue corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction).
+These contours in turn give a measure of the maximum amplification of a linear disturbance: the further negative contours extend into the upper-half plane, the larger a linear disturbance can grow even as a result of non-normality.
+This linear effect is captured entirely by GQL (and QL, for that matter).
+What role it plays in the development of subcritical spiral turbulence remains an open question.
+\begin{figure}
+    \centering
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_900.00_eta_0.8830_mu_-3.3333_nr_64_m_6_kz_ 9.68.pdf}
+    \end{minipage}
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \includegraphics[width=0.9\textwidth]{figs/TC_pseudospectrum_Re1_700.00_eta_0.8830_mu_-4.2857_nr_64_m_6_kz_ 9.68.pdf}
+    \end{minipage}
+    \caption{Spectra (blue points) and pseudospectra (contours) for $\Reyn_i = 900$ (left) and $\Reyn_i = 700$ (right) for a mode with $m=6$ and $k_z = 9.68$.
+The latter is the z wavenumber at the critical $\Reyn_{i,crit} \simeq 778$.
+$\omega$ is the frequency and $\gamma$ the growth rate at each point in the domain.
+The stability boundary is marked with a gray line at $\gamma = 0$.
+The are seven equally spaced contours from $-1.8$ to 0.}
+    \label{fig:pseudospectra}
+\end{figure}
 \section{Results}

### 2022-08-15 python/paths.py
updated KE vs time plots.

@@ -10,6 +10,16 @@ path2 = ["TC_3d_re_6.400000e+02_eta_8.830000e-01_mu_-1.875000e+00_Gamma_2.990000
          "TC_3d_re_5.200000e+02_eta_8.830000e-01_mu_-2.307692e+00_Gamma_2.990000e+01_M1_5_512_512_64"
          ]
 
+rei700_reo_3398 = ["TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_0_Lambdat_0",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_1_Lambdat_1",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_3_Lambdat_3",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_5_Lambdat_5",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_10_Lambdat_10",
+#                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_20_Lambdat_20",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_30_Lambdat_30",
+                   "TC_3d_re_7.000000e+02_eta_8.830000e-01_mu_-4.285714e+00_Gamma_2.990000e+01_M1_5_512_512_64_GQL_Lambdaz_60_Lambdat_60"
+]
 
[…]

### 2022-08-15 python/plot_KE_DNS_GQL.py

@@ -3,13 +3,12 @@ import h5py
 from series import RunSeries
 import matplotlib.pyplot as plt
 import numpy as np
-#from paths import rei640_reo_1359
-from paths import rei900_reo_3398
+import paths
 
 plt.style.use('prl')
-#path = RunSeries(rei640_reo_1359, 'rei640_reo_-1359')
-path = RunSeries(rei900_reo_3398, 'rei900_reo_-3398')
-
+#path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
+#path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
+path = RunSeries(paths.rei700_reo_3398, 'rei700_reo_-3398')
 filetype = Path("scalar/scalar.h5")
 path.data = []
 path.t = []
@@ -23,12 +22,15 @@ for i, r in enumerate(path.run_list):
     label = path.Lambdaz[i]
 
     if label is None:
-        label = '\infty'
+        label = 'DNS'
+    else:
+        label = f"$\Lambda = {label}$"
     print(f"Lambda = {label} number of samples = {path.data[i].shape[0]}")
-    plt.plot(path.t[i]/t_visc, path.data[i], label=f"$\Lambda = {label}$")
+    plt.plot(path.t[i]/t_visc, path.data[i], label=label)
 plt.xlabel(r"$t/\tau_\nu$")
 plt.ylabel(r"$E_{kin}$")
-plt.xlim(1,2.5)
-plt.legend()
+
+plt.xlim(3,4.5)
+plt.legend(ncol=2,fontsize=16)
 plt.tight_layout()
 plt.savefig(f"../figs/{path.name}_KE_vs_t.pdf")

### 2022-08-15 paper/paper.tex
Updates from Overleaf

@@
 %%%% Article title to be placed here
-\title{Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Flow}
+\title{Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Spiral Turbulence}
 
@@
 Some of the most common forms of turbulence in nature are anisotropic, with a symmetry broken by rotation, magnetic fields, or boundary effects.
-In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that contribute to the mean flow.
-Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively, discarding those that involve only high modes (see \cite{2022arXiv220505513M} for a recent review).
+In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that interact with the mean flow.
+Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review).
 GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.
@@
 In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}.
-We can always write the Navier-Stokes equations in schematic form as
+We can write the Navier-Stokes equations in schematic form as
 \begin{equation}
@@
 \end{equation}
-where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u}  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
-As noted by \cite{2002PhFl...14.3475H}, the linear operator $\Lop$ for TC flow is a function of its control parameters.
+where $\Lop$ is a linear operator and $\mathcal{N}$ is a non-linear function of the state vector $\mathbf{q} = \left[ \mathbf{u},  p\right]^T$, where $\mathbf{u}$ is the velocity field and $p$ the pressure.
+As noted by \cite{2002PhFl...14.3475H}, the non-normality of the linear operator $\Lop$ for TC flow is a function of its control parameters.
 That is, its commutator with its adjoint
@@
 We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalization, and numerical methods.
-Section~\ref{sec:linear} reviews both modal and non-modal linear analyses and contains a new analysis combining linear and non-linear aspects.
-In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL in each of the three regimes.
+Section~\ref{sec:linear} reviews both modal and non-modal linear analyses of spiral turbulence.
+In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL for supercritical and subcritical spiral turbulence.
 Finally, we offer discussion and concluding results in section~\ref{sec:conclusion}.
@@
 We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, most unstable mode is never axisymmetric.
-At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset, and
+At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset.
+Figure~\ref{fig:LSB} also shows the hysteresis boundaries given in references \cite{1986JFM...164..155A} and \cite{1965JFM....21..385C}.
+In both cases, original tabular data was not available, so we extracted the data from figures 16 and 2a, respectively using WebPlotDigitizer \cite{Rohatgi2020}.
+Note the error bars from Coles (1965) \cite{1965JFM....21..385C} give the top and bottom of the markers from the original paper.
+Notably, even with this uncertainty, the two studies do not agree on the lower boundary for hysteresis despite similar experimental geometries.
 \begin{figure}
@@
 Solid circles shows points where we have conducted GQL analysis; crosses represent DNS steps along the path to subcritical behaviour.
-}
+The curves labelled Andereck et al.
+(1986) and Coles (1965) give the boundaries of spiral turbulence hysteresis reported in those two papers.}
     \label{fig:LSB}
@@
 
-Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ 
+Following pioneering numerical work \cite{2009PhRvE..79c6309M, 2009PhRvE..80d6315M}, we set the outer cylinder rotation rate to $\Reyn_o = -3398$ and
 %for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and 
@@
 \begin{equation}
+\label{eqn:pseudospectrum}
     \lVert (\sigma \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
@@
 for a given complex number $\sigma = \gamma + i\omega$ here representing the eigenvalue corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction).
-These contours in turn give a measure of the maximum amplification of a linear disturbance: the further negative contours extend into the upper-half plane, the larger a linear disturbance can grow even as a result of non-normality.
+Equation~\ref{eqn:pseudospectrum} depends on the choice of norm; here we use the standard energy norm.
+These contours in turn give a measure of the maximum amplification of a linear disturbance: the further negative contours extend into the upper-half plane, the larger a linear disturbance can grow as a result of non-normality, irrespective of the presence of instability.
 This linear effect is captured entirely by GQL (and QL, for that matter).
@@
 Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the rms velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$.
-Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$.
+Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$ (not shown).
 \begin{figure}
@@
 \end{figure}
-
+Finally, in figure~\ref{fig:rei900_vmean}, we show the azimuthal velocity averaged over $theta$ and $z$, giving the mean deviation from CCF as a function of GQL cutoff $\Lambda$.
+Note that no conditional averaging \cite{2009PhRvE..80f7301D} is performed here, so this represents an average over both laminar and turbulent regions for DNS.
+As $\Lambda$ is increased, the profile approaches the DNS case; this is typical of GQL behaviour (figure 7 of \cite{2017JFM...810..412T} shows the same for rotating plane Couette flow).
 \begin{figure}
@@
 Understanding how these interactions work to shape subcritical TC turbulence represents a promising new avenue to understanding how ECS are maintained and how they in turn sustain turbulence below linear stability bounds.
-In particular, one can begin to understand the relative importance of
+In particular, one can begin to understand the relative importance of non-normal effects (represented entirely with GQL) and non-linear ECS.
 

### 2022-08-15 python/plot_v_r_DNS_GQL.py
updated v profiles.

@@ -3,29 +3,48 @@ import h5py
 from series import RunSeries
 import matplotlib.pyplot as plt
 import numpy as np
-from paths import rei640_reo_1359
-
+import paths
 plt.style.use('prl')
-path = RunSeries(rei640_reo_1359, 'rei640_reo_-1359')
+#path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
+path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
 
-filetype = Path("slices/slices_s1.h5")
+t_cutoff = 1.7
+filetype = Path("slices/slices.h5")
 path.data = []
 path.r = []
-
-for i, r in enumerate(path.run_list):
+path.sim_time = []
+colors = plt.rcParams["axes.prop_cycle"].by_key()["color"]
+for i, r in enumerate(path.run_list[:-2]):
     filename = path.root_dir/Path(r)/filetype
     with h5py.File(filename, "r") as df:
         path.data.append(df['tasks/v_tot'][:,0,0,:])
         path.r.append(df['scales/r/1.0'][:])
+        path.sim_time.append(df['scales/sim_time'][:])
     label = path.Lambdaz[i]
-
+    eta = path.eta[i]
+    mu = path.mu[i]
+    t_visc = 1/path.Re[i]
+    A = (1/eta - 1.)*(mu-eta**2)/(1-eta**2)
+    B = eta*(1-mu)/((1-eta)*(1-eta**2))
+    r = path.r[i]
+    vr = A*r + B/r
+    dns_profile = path.data[0][path.sim_time[0]/t_visc>t_cutoff,:].mean(axis=0)
+    r_profile = path.data[i][path.sim_time[i]/t_visc>t_cutoff,:].mean(axis=0)
     if label is None:
-        label = '\infty'
+        label = 'DNS'
+        c = 'k'
+        lw = 3
+    else:
+        label = f"$\Lambda = {label}$"
+        c = colors[i]
+        lw = 2
+        l2_norm = ((r_profile - dns_profile)**2).mean()
+        print(f"Lambda = {label} l2_norm = {l2_norm}")
     print(f"Lambda = {label} number of samples = {path.data[i].shape[0]}")
-    plt.plot(path.r[i],path.data[i][-25:,:].mean(axis=0), label=f"$\Lambda = {label}$")
[…]

### 2022-08-15 python/plot_rei900_snapshots_figure.py
updated snapshots figure

@@ -11,7 +11,7 @@ plt.style.use('prl')
 #path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
 path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
 
-aspect=0.9 #np.sqrt(2)
+aspect=1.2 #np.sqrt(2)
 w = 5
 h = aspect*w
 
@@ -25,10 +25,22 @@ def load_data(df, Rmid, index=-1):
 
     return xmesh, ymesh, dset
 
+def load_snapshot_data(df, Rmid, index=-1):
+    mid = int(df['tasks']['u'].shape[-1]/2)
+    dset = np.sqrt(df['tasks']['u'][index, :,:,mid]**2 + df['tasks']['v'][index, :,:,mid]**2 + df['tasks']['w'][index, :,:,mid]**2)
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),mid]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/u'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+
+
 
 fig = plt.figure(figsize=(w, h))
 grid = AxesGrid(fig, 111,
-                nrows_ncols=(3, 2),
+                nrows_ncols=(4, 2),
                 axes_pad=0.05,
                 cbar_mode='single',
                 cbar_location='top',
@@ -37,7 +49,7 @@ grid = AxesGrid(fig, 111,
 
 
 for i,ax in enumerate(grid):
-    if i != 4:
+    if i != 6:
         ax.set_axis_off()
     else:
         ax.set_xlabel(r"$R_{mid}\theta$", fontsize=14)
@@ -50,11 +62,15 @@ for i,ax in enumerate(grid):
     filetype = 'slices/slices_s2.h5'
     if not (path.root_dir/Path(r)/filetype).exists():
         filetype = 'slices/slices_s1.h5'
+    if not (path.root_dir/Path(r)/filetype).exists():
+        filetype = 'snapshots/snapshots_s4.h5'
     filename = path.root_dir/Path(r)/filetype
-
     Rmid = eta/(1-eta) + 0.5
     with h5py.File(filename, "r") as df:
-        xgrid, ygrid, data = load_data(df, Rmid)
+        try:
+            xgrid, ygrid, data = load_data(df, Rmid)
+        except:
[…]

### 2022-08-15 python/plot_rei900_spectra_figure.py
added spectra figure.

@@ -1,0 +1,107 @@
+from pathlib import Path
+import h5py
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import AxesGrid
+import numpy as np
+import paths
+from series import RunSeries
+from dedalus.extras import plot_tools
+plt.style.use('prl')
+
+#path = RunSeries(paths.rei640_reo_1359, 'rei640_reo_-1359')
+path = RunSeries(paths.rei900_reo_3398, 'rei900_reo_-3398')
+
+index = -1 # time index
+eps = 1e-16
+aspect=1.6 #np.sqrt(2)
+w = 5
+h = aspect*w
+nk = 256
+nm = 256
+
+def load_data(df, Rmid, index=-1):
+    dset = df['tasks']['v_r_slice'][index, :,:,0]
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),0]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/u_r_slice'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+def load_snapshot_data(df, Rmid, index=-1):
+    mid = int(df['tasks']['v'].shape[-1]/2)
+    dset = df['tasks']['v'][index, :,:,mid]
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),mid]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/v'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+
+def load_data(df, Rmid, index=-1):
+    dset = df['tasks']['v_r_slice'][index, :,:,0]
+
+    image_axes = [2, 1]
+    data_slices = [index, slice(None), slice(None),0]
+    xmesh, ymesh, vdata = plot_tools.get_plane(df['tasks/u_r_slice'], 2, 1, data_slices)
+    xmesh *= Rmid
+
+    return xmesh, ymesh, dset
+
+
+fig = plt.figure(figsize=(w, h))
+grid = AxesGrid(fig, 111,
+                nrows_ncols=(4, 2),
+                axes_pad=0.05,
+                cbar_mode='single',
+                cbar_location='top',
+                cbar_pad=0.
+                )
+
+
+for i,ax in enumerate(grid):
+    if i != 6:
+        ax.set_axis_off()
+    else:
[…]

### 2022-08-15 paper/paper.tex

@@
     \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$.
-On the left,  $\Reyn_i = 900$ is above the linear stability boundary, while the right shows $\Reyn_i = 700$, below the stability boundary.
-Starting from a saturated spiral turbulence state at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10$.
-For $\Reyn_i = 700$, we also use $\Lambda = 20, 30$ At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations.
+$\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary.
+$\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary.
+Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$.
+At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations.
+At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state.
 }
@@
 \end{figure}
-Finally, in figure~\ref{fig:rei900_vmean}, we show the azimuthal velocity averaged over $theta$ and $z$, giving the mean deviation from CCF as a function of GQL cutoff $\Lambda$.
+Finally, in figure~\ref{fig:rei900_vmean}, we show the azimuthal velocity averaged over $\theta$ and $z$, giving the mean deviation from CCF as a function of GQL cutoff $\Lambda$.
 Note that no conditional averaging \cite{2009PhRvE..80f7301D} is performed here, so this represents an average over both laminar and turbulent regions for DNS.
 As $\Lambda$ is increased, the profile approaches the DNS case; this is typical of GQL behaviour (figure 7 of \cite{2017JFM...810..412T} shows the same for rotating plane Couette flow).
+However, unusually, QL performs slightly better than low order GQL with $\Lambda < 5$.
 \begin{figure}
@@
 Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study.
-HThe right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$.
-Our very preliminary data for $\Lambda = 30$ show that it recovers from the decay and begins to climb in energy again, similar to the early behaviour from the supercritical runs.\footnote{\textbf{Note to reviewers}: this simulation is still running as of the 31 July due date for the initial manuscript.
-By the time refereeing begins, we will update this plot and section.}
+The right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$.
+At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again, peaking in a state .
 

### 2022-08-15 python/plot_v_r_DNS_GQL.py
fixed vmean profile...

@@ -41,7 +41,7 @@ for i, r in enumerate(path.run_list[:-2]):
         l2_norm = ((r_profile - dns_profile)**2).mean()
         print(f"Lambda = {label} l2_norm = {l2_norm}")
     print(f"Lambda = {label} number of samples = {path.data[i].shape[0]}")
-    plt.plot(r, - vr, label=label, color=c, lw=lw)
+    plt.plot(r, r_profile -vr, label=label, color=c, lw=lw)
 plt.xlabel(r"$r/d$")
 plt.ylabel(r"$\left< v' \right>$")
 plt.xlim(path.r[0].min(),path.r[0].max())

### 2022-08-15 paper/paper.tex
Updates from Overleaf

@@
 \end{figure}
+
 \subsection{Supercritical Spiral Turbulence}
 We first consider the supercritical case with $\Reyn_i = 900$.
-We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
+We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10, 30, 60$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
 Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$.
-The first important thing to note is that none of these low-order models are able to maintain the spiral structure.
-However, even including a single additional low mode, $\Lambda = 1$, shows a significant improvement in flow morphology compared to QL.
+The first important thing to note is that low-order models are not able to maintain the spiral structure until $\Lambda \geq 30$.
+However, even including a single additional low mode, $\Lambda = 1$, shows a significant increase in flow complexity compared to QL.
 It is quite interesting to note that QL \emph{does} feature some degree of spatial intermittency--there are clearly separated regions hosting spiral waves interspersed with featureless, laminar patches.
@@
 The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$.
-Even at $\Lambda = 10$, spiral turbulence is not apparent.}
+Until $\Lambda > 30$, spiral turbulence is not apparent.}
     \label{fig:rei900_snapshots}
@@
 Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
-The supercritical case shows significant oscillations at all $\Lambda$. 
+The supercritical case shows significant oscillations at all $\Lambda < 30$.
+While both $\Lambda 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
+Interestingly, $\Lambda=30$ has a lower kinetic energy much closer to that of DNS than does $\Lambda=60$.
+By examining power spectra (figure
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
+    \caption{$\theta$-$z$ power spectra of $v$ at midgap for $\Reyn_i = 900$ at various GQL cutoffs.
+Adding a single GQL mode significantly improves the power spectrum.
+Spiral turbulence appears at $\Lambda=30$.
+The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
+    \label{fig:my_label}
+\end{figure}
 Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g.
@@
 As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT).
-Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches is closests approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.
+Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches its closest approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.
 
@@
 As $\Lambda$ is increased, the profile approaches the DNS case; this is typical of GQL behaviour (figure 7 of \cite{2017JFM...810..412T} shows the same for rotating plane Couette flow).
-However, unusually, QL performs slightly better than low order GQL with $\Lambda < 5$.
+However, QL performs slightly better than low order GQL with $\Lambda < 5$.
 \begin{figure}
@@
 We have presented a series of simulations demonstrating the performance of GQL at various cutoffs for Taylor-Couette flow in the spiral turbulence regime.
-GQL is a significant improvement over QL models in representing mean quantities, including the deviation from circular Couette flow.
-However, even at the highest cutoffs studied here, it does not retain an unambiguous spiral form in either the sub- or supercritical regimes.
+GQL is a significant improvement over QL models in representing statistical quantities, particularly the power spectra.
+However, it does not retain an unambiguous spiral form in the supercritical regime until $\Lambda \simeq 30$; at subcritical $\Reyn_i$, $\Lambda \simeq 60$ is required to see any sustained flows at all.
+Interestingly, when GQL solutions do sustain non-CCF flows in the subcritical regime, they also produce spiral patterns.
 Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}.
 This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realizations \cite{2022arXiv220505513M}.
-
-However, we have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow.
+We have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow.
 In future work, we will consider the ability of GQL to maintain exact coherent states identified by Deguchi et al \cite{2014PhRvL.112r4502D} and the dynamics of which were studied recently by Wang and coworkers \cite{2022arXiv220712990W}.

### 2022-08-15 python/plot_rei900_spectra_figure.py
zoomed in spectra.

@@ -16,8 +16,8 @@ eps = 1e-16
 aspect=1.6 #np.sqrt(2)
 w = 5
 h = aspect*w
-nk = 256
-nm = 256
+nk = 100
+nm = 100
 
 def load_data(df, Rmid, index=-1):
     dset = df['tasks']['v_r_slice'][index, :,:,0]

### 2022-08-15 paper/paper.tex

@@
 The supercritical case shows significant oscillations at all $\Lambda < 30$.
-While both $\Lambda 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
-Interestingly, $\Lambda=30$ has a lower kinetic energy much closer to that of DNS than does $\Lambda=60$.
-By examining power spectra (figure
+While both $\Lambda = 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
+Interestingly, figure~\ref{fig:ke_vs_t_rei900} shows that $\Lambda=30$ has a \emph{lower} kinetic energy much closer to that of DNS than does $\Lambda=60$.
+By examining power spectra in figure~\ref{fig:rei900_spectra}, we begin to see the reason for the more complex flows seen with $\Lambda > 0$.
+Even low-order non-linear scatter leads to a steady increase
 \begin{figure}
@@
 The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
-    \label{fig:my_label}
+    \label{fig:rei900_spectra}
 \end{figure}
@@
 The right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$.
-At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again, peaking in a state .
+At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again but ultimately returns to the laminar CCF state.
+Curiously, our $\Lambda = 60$ run not only remains in a non-laminar solution for as long as we have been able to run for, it also correctly reproduces the spiral turbulence pattern, just as in the supercritical case.
+
+The ability of GQL to reproduce spiral turbulence in the subcritical regime is quite significant.
+First, while $\Lambda = 60$ corresponds to the fully non-linear interactions occurring in a rather large fraction of the total number of modes in the simulation, it is important to note that a DNS at $n_\theta = n_z = 60$ would be massively underresolved, possibly numerically unstable.
+Second, the ability of non-linear interactions to sustain turbulence where non-normal linear dynamics alone cannot suggests an important role for exact coherent states in subcritical Taylor-Couette flow.
+Because GQL retains all linear dynamics and low-order non-linear couplings, in addition to the couplings between the mean flows and high modes, if non-normal dynamics played a strong role in subcritical turbulence, one would expect the behavior seen at $\Lambda = 30$ to lead to sustained, rather than transient turbulence.
 

### 2022-08-15 python/plot_rei900_spectra_figure.py
fixed spectra labels.

@@ -93,7 +93,7 @@ for i,ax in enumerate(grid):
     label = path.Lambdaz[i]
     if label is None:
         label = '\infty'
-    ax.text(180,225,f"$\Lambda = {label}$", color='white')
+    ax.text(70,80,f"$\Lambda = {label}$", color='white')
 
 # when cbar_mode is 'single', for ax in grid, ax.cax = grid.cbar_axes[0]
 

### 2022-08-16 paper/paper.tex
Updates from Overleaf

@@
 This flow state is quite complex, with delicate internal structure, and it can be traced into certain regimes of linear stability.
-This behavior is believed to be connected to the non-normality of the linear operator, which is itself a function of the control parameters.
-Using spiral turbulence in both linearly stable and unstable regimes, we investigate the effectiveness of the generalized quasilinear approximation (GQL), an extension of quasilinear theory designed to capture the essential aspects of turbulent flows.
+This behaviour is believed to be connected to the non-normality of the linear operator, which is itself a function of the control parameters.
+Using spiral turbulence in both linearly stable and unstable regimes, we investigate the effectiveness of the generalised quasilinear approximation (GQL), an extension of quasilinear theory designed to capture the essential aspects of turbulent flows.
 We find that GQL performs much better in the supercritical regime than the subcritical.
@@
 In these situations, significant theoretical progress can be made by considering \emph{quasi-linear} (QL) models, in which one retains only those non-linearities that interact with the mean flow.
-Recently, the generalized quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review).
+Recently, the generalised quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review).
 GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.
@@
 Here, we consider Taylor-Couette (TC) flows at moderate inner and outer Reynolds numbers, $|\Reyn_{i,o}| < 3500$.
-This is an interesting testbed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$.
+This is an interesting test bed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$.
 One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
@@
 More recent work has identified both supercritical and subcritical paths to spiral turbulence\cite{2009PhRvE..80d6315M}.
-In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarizes below the stability boundary when $\Reyn_o = -1200$.
+In that work, the authors follow spiral turbulence along a decreasing $\Reyn_i$ path and show that the state is maintained even below the linear stability boundary when $\Reyn_o = -3000$ but it relaminarises below the stability boundary when $\Reyn_o = -1200$.
 Using \cite{2009PhRvE..80d6315M} as a guide, we use direct numerical simulations (DNS) at $\Reyn_o = -3398$ and decrease $\Reyn_i$ into the subcritical region.
@@
 
-We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalization, and numerical methods.
+We begin in section~\ref{sec:methods} by detailing our equations, non-dimensionalisation, and numerical methods.
 Section~\ref{sec:linear} reviews both modal and non-modal linear analyses of spiral turbulence.
@@
 \end{equation}
-It is also quite useful to characterize times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
+It is also quite useful to characterise times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
 
@@
 This linear effect is captured entirely by GQL (and QL, for that matter).
-What role it plays in the development of subcritical spiral turbulence remains an open question.
+What role it plays in the development of subcritical spiral turbulence remains an open question, though our results provide some interesting constraints.
 \begin{figure}
@@
 $\omega$ is the frequency and $\gamma$ the growth rate at each point in the domain.
-The stability boundary is marked with a gray line at $\gamma = 0$.
+The stability boundary is marked with a grey line at $\gamma = 0$.
 The are seven equally spaced contours from $-1.8$ to 0.}
@@
 \label{sec:nonlinear}
-Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the rms velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$.
+Figure~\ref{fig:urms_tz_rei700} shows a typical DNS snapshot of the RMS velocity perturbation $u'_{rms}$ in the middle of the gap for $\Reyn_i=700$.
 Despite being well below the linear stability threshold $\Reyn_{i,crit} \simeq 778$, the flow morphology is nearly identical to the supercritical case $\Reyn_i = 900$ (not shown).
@@
     \includegraphics[width=\textwidth]{../figs/urms_tz_rei_700_reo-3398_000278.png}
-    \caption{Slice of fluctuating rms velocity in the middle of the radial domain at $\Reyn_i = 700$, $\Reyn_o = -3398$. This solution is well below the linear stability threshold at }
+    \caption{Slice of fluctuating RMS velocity in the middle of the radial domain of a DNS at $\Reyn_i = 700$, $\Reyn_o = -3398$.
+This solution is well below the linear stability threshold at $\Reyn_{i,crit} \simeq 778$.}
     \label{fig:urms_tz_rei700}
@@
 We first consider the supercritical case with $\Reyn_i = 900$.
-We initialize GQL simulations with $\Lambda = 0, 1, 3, 5, 10, 30, 60$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
+We initialise GQL simulations with $\Lambda = 0, 1, 3, 5, 10, 30, 60$ using a snapshot of DNS data at $t \simeq 1.5 \tau_\nu$.
 Figure~\ref{fig:rei900_snapshots} shows slices at mid-gap, $r = R_{mid}$.
@@
     \caption{RMS velocity perturbations in the $z-\theta$ plane at $r=R_{mid}$ for $\Reyn_i=900, \Reyn_0=-3398$.
-The upper left panel shows DNS, the upper right panel shows QL, and the remaining four are GQL with $\Lambda = 1,3,5,10$.
+The upper left panel shows DNS, the upper right panel shows QL, and the remaining six are GQL with $\Lambda = 1,3,5,10, 30, 60$.
 Until $\Lambda > 30$, spiral turbulence is not apparent.}
@@
 Interestingly, figure~\ref{fig:ke_vs_t_rei900} shows that $\Lambda=30$ has a \emph{lower} kinetic energy much closer to that of DNS than does $\Lambda=60$.
+
 By examining power spectra in figure~\ref{fig:rei900_spectra}, we begin to see the reason for the more complex flows seen with $\Lambda > 0$.
-Even low-order non-linear scatter leads to a steady increase
+Even low-order non-linear scatter leads to a steady increase to higher $k$ and $m$.
+The spectra of $\Lambda = 30$ and 60 are of particular interest in understanding the nature of spiral turbulence: they show a transition toward filamentary structure aligned along the diagonal, very similar to DNS.
+The square feature clearly visible at $\Lambda = 60$ is analogous to similar structures seen in channel flow turbulence (see figure 4.20 of \cite{2019Kellam}).
+In that work, the author proposed a potential explanation in terms of the linear instability of the non-linear solution: assume there is an instability of the mean $(m, k_z) = (0,0)$ at some wavenumber $k'$.
+If $k'$ is counted among the high modes, high-high $\to$ high interactions smoothly populate the high modes.
+However, as $\Lambda$ increases, it reaches a point at which the last unstable mode $k'$ becomes counted among the low modes.
+Because low-low $\to$ high interactions are truncated in GQL,  this results in a sudden decrease in the amount of energy available in the high part of the spectrum.
+Reference \cite{2019Kellam} constructs a preliminary model of this behaviour for plane channel flow and finds support for the idea.
 \begin{figure}
@@
     \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
-    \caption{$\theta$-$z$ power spectra of $v$ at midgap for $\Reyn_i = 900$ at various GQL cutoffs.
+    \caption{$\theta$-$z$ power spectra of $v$ at mid-gap for $\Reyn_i = 900$ at various GQL cutoffs.
 Adding a single GQL mode significantly improves the power spectrum.
 Spiral turbulence appears at $\Lambda=30$.
-The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
+The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.
+The figure shows only the lowest 100 modes in each direction, though in the simulations $k_{max} = m_{max} = 512$.
+}
     \label{fig:rei900_spectra}
@@
 Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}.
-This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realizations \cite{2022arXiv220505513M}.
+This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realisations \cite{2022arXiv220505513M}.
 We have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow.

### 2022-10-10 python/TC_pseudospectrum.py
add capability to compute pseudospectra for saturated backgrounds.

@@ -1,6 +1,6 @@
 """
 Usage:
-  TC_pseudospectrum.py [--re=<reynolds>] [--eta=<eta>] [--m=<initial_m>] [--kz=<kz>] [--mu=<mu>] [--nr=<nr>] 
+  TC_pseudospectrum.py [--re=<reynolds>] [--eta=<eta>] [--m=<initial_m>] [--kz=<kz>] [--mu=<mu>] [--nr=<nr>] [<filename>]
 
 Options:
   --re=<reynolds>  Reynolds number for simulation [default: 250]
@@ -26,6 +26,7 @@ import numpy as np
 from dedalus import public as de
 from eigentools import Eigenproblem, CriticalFinder
 import matplotlib.pyplot as plt
+import h5py
 
 args=docopt(__doc__)
 Re1=float(args['--re'])
@@ -34,8 +35,7 @@ mu = float(args['--mu'])
 nr=int(args['--nr'])
 m = int(args['--m'])
 alpha = float(args['--kz'])
-
-filename = "../figs/TC_pseudospectrum_Re1_{:5.2f}_eta_{:6.4f}_mu_{:6.4f}_nr_{:d}_m_{:d}_kz_{:5.2f}.pdf".format(Re1, eta, mu, nr, m, alpha)
+infile = args['<filename>']
 
 def energy_norm(Q1, Q2):
     u1 = Q1['u']
@@ -94,12 +94,29 @@ Lap_t --> theta component of vector laplacian
 Lap_z --> z component of vector laplacian
 
 """
+outfilename = "../figs/TC_pseudospectrum_Re1_{:5.2f}_eta_{:.4f}_mu_{:.4f}_nr_{:d}_m_{:d}_kz_{:.2f}".format(Re1, eta, mu, nr, m, alpha)
 
-problem.substitutions['A'] = '(1/eta - 1.)*(mu-eta**2)/(1-eta**2)'
-problem.substitutions['B'] = 'eta*(1-mu)/((1-eta)*(1-eta**2))'
 
-problem.substitutions['v0'] = 'A*r + B/r'       #background profile? forcing instead of forcing the boundaries
-problem.substitutions['dv0dr'] = 'A - B/(r*r)'  #d/dr of background forcing
+if infile:
+    print(f"Running with background from {infile}")
+    V = domain.new_field()
+    dVdr = domain.new_field()
+    V.meta['r']['constant'] = True
+    dVdr.meta['r']['constant'] = True
[…]

### 2022-10-11 python/plot_KE_DNS_GQL.py
updated KE figs for referees.

@@ -30,7 +30,11 @@ for i, r in enumerate(path.run_list):
 plt.xlabel(r"$t/\tau_\nu$")
 plt.ylabel(r"$E_{kin}$")
 
-plt.xlim(3,4.5)
-plt.legend(ncol=2,fontsize=16)
+plt.xlim(3.4,4.4)# for Rei=700
+#plt.xlim(1.2,2.4) # for Rei=900
+plt.ylim(0,420) # for Rei=700
+#plt.ylim(0,700) # for Rei=900
+plt.legend(ncol=2,fontsize=14) # for Rei=700
+#plt.legend(ncol=1,fontsize=14) # for Rei=900
 plt.tight_layout()
 plt.savefig(f"../figs/{path.name}_KE_vs_t.pdf")

### 2022-10-11 paper/paper.tex
Updates from Overleaf

@@
 \label{sec:intro}
-%Something about Taylor (1923); GQL.
 Since Taylor's classic 1923 paper\cite{1923RSPTA.223..289T}, fluid between two independently rotating cylinders has become a rich laboratory for the understanding of instability, transition to turbulence, and turbulence itself.
@@
 Recently, the generalised quasi-linear approximation (GQL) has been developed to systematically expand the QL idea to include sets of non-linear interactions between two \emph{sets} of modes, low and high, respectively (see \cite{2022arXiv220505513M} for a recent review).
-GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam}.
+GQL has been shown to perform significantly better in reproducing direct numerical simulations (DNS) than QL models across a wide range of paradigm flows including forced zonal jet formation \cite{2016PhRvL.116u4501M}, rotating plane Couette flow \cite{2017JFM...810..412T}, rapidly rotating convection \cite{2018RSPSA.47480422T}, and channel flow\cite{2019Kellam, 2022JFM...936A..33H, 2022JFM...944A..34H}.
 
@@
 This is an interesting test bed for GQL because it features numerous intricate patterns that are maintained by non-linear interactions beyond the cascade, local in coefficient space, found at high $\Reyn$.
-One of the key feature of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
+One of the key features of TC flow is that it has non-linear and linear instabilities in different regions of parameter space.
 In the linearly stable regions, transient growth has been observed in TC flow\cite{2002PhFl...14.3475H, 2002PhFl...14.1655M}.
@@
 \end{equation}
-can be zero or non-zero. 
+can be zero or non-zero, and $f$ is a function of the four parameters of Taylor-Couette flow: the inner cylinder Reynolds number $\Reyn_i$, the ratio of inner to outer radii $\eta \equiv R_i/R_o$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_o/\Omega_i$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$.
 This provides a very useful arena to test the validity of GQL in a system that admits a varying degree of non-normality.
@@
 In section~\ref{sec:nonlinear}, we describe the main results of the paper, the comparison of DNS, QL, and GQL for supercritical and subcritical spiral turbulence.
-Finally, we offer discussion and concluding results in section~\ref{sec:conclusion}.
+Finally, we offer discussion and concluding remarks in section~\ref{sec:conclusion}.
 %Recently, a subcritical transition to turbulence via interpenetrating spirals has been reported \cite{2020JFM...892A..12C}.
@@
 We solve the incompressible Navier-Stokes equations with unit density for $\mathbf{u}' = \mathbf{u} - \mathbf{U}$,
-where $\mathbf{U} = A r + B/r \hat{\mathbf{e}}_\phi$ is circular Couette flow (CCF),
+where $\mathbf{U} = A r + B/r \hat{\mathbf{e}}_\phi$ is circular Couette flow (CCF) and $A$ and $B$ are the standard constants (see \cite{Chandra1961}, chapter 7),
 \begin{align}
@@
       \frac{\partial \mathbf{u'}}{\partial t} + \mathbf{U} \cdot \nabla \mathbf{u'} + \mathbf{u'} \cdot \nabla \mathbf{U} + \mathbf{u'} \cdot \nabla \mathbf{u'} &= -\nabla p + \nu \nabla^2\mathbf{u'}\\
-      \nabla \cdot \mathbf{u'} &= 0.
+      \nabla \cdot \mathbf{u'} &= 0,
 \end{align}
-In order to implement the GQL approximation, we introduce a projection operator to divide $\mathbf{u'}$ into two subsets,
+where $p$ is the pressure and $\nu$ is the kinematic viscosity.
+We solve equations~\ref{eqn:NS} without modification for the DNS runs.
+To briefly review, GQL extends the quaslinear approximation by dividing
+ $\mathbf{u'}$ into two subsets,
 \begin{align}
     \mathbf{u'}_l = \left< \mathbf{u'}\right>_l &= \sum_{m = -\Lambda_\theta}^{\Lambda_\theta} \sum_{k = -\Lambda_z}^{\Lambda_z} \mathbf{\hat{u}'}_{mk}(r) e^{i (m \theta + 2\pi k z/L_z)}\\
-    \mathbf{u'}_h = \mathbf{u'} - \mathbf{u'}_l.
+    \mathbf{u'}_h = \mathbf{u'} - \mathbf{u'}_l,
 \end{align}
-GQL can then be written as 
+which we refer to as the ``low'' and ``high'' modes, respectively.
+$\Lambda_z$ and $\Lambda_\theta$ are adjustable parameters that represent the \emph{cutoff} in wavenumber space delineating low from high. 
+The non-linearity, when expressed in spectral space is a triadic interaction between wavenumbers.
+GQL includes all such interactions that fall into three classes: low-low $\to$ low, high-high $\to$ low, and low-high $\to$ high, and it discards all others.
+This is a closed system that respects all conservation laws of the original equations \cite{2016PhRvL.116u4501M}.
+Most importantly, it represents a systematic closure of increasing fidelity independent of a particular theoretical model for the underlying turbulence.
+
+We introduce a projection operator to perform this division, in which case
+GQL can be written as 
 \begin{align}
@@
 This form is equivalent to those given in prior works (for example \cite{2017JFM...810..412T}), where separate equations are written for the low and high modes, respectively.
-For all runs in this paper, $\Lambda = \Lambda_z = \Lambda_\theta$, though we note it would be interesting to consider models quasilinear in $\theta$ (i.e.
-$\Lambda_\theta = 0$) with varying $\Lambda_z$.
-These correspond to the \emph{restricted non-linear} models used with success in shear flow turbulence \cite{2014PhFl...26j5112T, 2015PhFl...27j5104T, 2015PhFl...27a1702B}.
 
 We non-dimensionalise lengths in terms of the gap width $d = R_2 - R_1$ and velocities in terms of the inner cylinder velocity $U_i = R_i \Omega_i$.
-Taylor-Couette flow can be parameterised in terms of four parameters, the inner cylinder Reynolds number $\Reyn_i$, the ratio of inner to outer radii $\eta \equiv R_i/R_o$, the ratio of outer to inner rotation rates $\mu \equiv \Omega_o/\Omega_i$, and the aspect ratio of the cylinder $\Gamma \equiv L_z/d$.
 In our computations, the $z$ direction is periodic and thus $L_z$ is the periodicity length.
@@
 It is also quite useful to characterise times in terms of the viscous diffusion time $\tau_\nu \equiv d^2/\nu$. In our units, $\tau_\nu = \Reyn_i$.
+
+In our simulations, the possible spiral patterns are quantised by the periodicity in $z$ and $\theta$.
+Figure~\ref{fig:urms_tz_rei700} shows that for $\Reyn_i = 700$, the pattern selected has $k_z = m = 1$; $\Reyn_i = 900$ has the same pattern (not shown).
+Given this pattern, it is unclear a priori how to select the cutoff wavenumbers: unlike the case of plane Couette flow, where a self-sustaining process involving a quasilinear transfer of streamwise waves to spanwise/wall-normal rolls is well established, here we have a turbulent structure that is bounded in both $z$ and $\theta$.
+Given this lack of apparent directionality, for all runs in this paper, $\Lambda = \Lambda_z = \Lambda_\theta$, though we note it would be interesting to consider models with varying $\Lambda_\theta$  but $\Lambda_z = N_z$, where $N_z$ is the DNS resolution.
+Such models are fully non-linear in $z$ and thus generalise the quasilinear \emph{restricted non-linear} models used with success in shear flow turbulence \cite{2014PhFl...26j5112T, 2015PhFl...27j5104T, 2015PhFl...27a1702B}.
 
@@
 900 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10 & Supercritical SPT\\
-700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30 & Subcritical SPT\\
+700 & -3398 & 0.883 & 29.9 & 64 & 512 & 512 & 0, 1, 3, 5, 10, 20, 30, 60 & Subcritical SPT\\
 %\hline
@@
 In order to determine the linear stability bound, we find the minimum $\Reyn_i$ for a given $\Reyn_o$ by solving the eigenvalue problem for equations~\ref{eqn:NS}.
-We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, most unstable mode is never axisymmetric.
-At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 5$ at onset.
+We note that for all $\Reyn_o$ plotted in figure~\ref{fig:LSB}, the most unstable mode is never axisymmetric.
+At $\Reyn_o=-3398$, where our non-linear calculations are run, the most unstable mode has $m = 6$ at onset.
 Figure~\ref{fig:LSB} also shows the hysteresis boundaries given in references \cite{1986JFM...164..155A} and \cite{1965JFM....21..385C}.
 In both cases, original tabular data was not available, so we extracted the data from figures 16 and 2a, respectively using WebPlotDigitizer \cite{Rohatgi2020}.
-Note the error bars from Coles (1965) \cite{1965JFM....21..385C} give the top and bottom of the markers from the original paper.
-Notably, even with this uncertainty, the two studies do not agree on the lower boundary for hysteresis despite similar experimental geometries.
+The error bars on the Coles (1965) \cite{1965JFM....21..385C} data give the top and bottom of the markers from the original paper.
+We note in passing that the two studies do not agree on the lower boundary for hysteresis despite similar experimental geometries.
+Our subcritical experiment is within the more conservative hysteresis boundary given by reference \cite{1986JFM...164..155A}.
 \begin{figure}
@@
 %for the subcritical path and $\Reyn_o = -1359$ for the supercritical path and 
-start a series of DNS runs at a $\Reyn_i$ high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random inital condition satisfying $\nabla \cdot \mathbf{u} = 0$.
-From these two seed runs, we decrease $\Reyn_i$ in steps shown in figure~\ref{fig:LSB} where each point represents a DNS or GQL simulation run for one viscous time.
+start a DNS run at $\Reyn_i = 900$, high enough to trigger spiral turbulence driven by linear instability from a low-amplitude, random initial condition satisfying $\nabla \cdot \mathbf{u} = 0$.
+From this seed run, we decrease $\Reyn_i$ in steps shown in figure~\ref{fig:LSB} where each point represents a DNS or GQL simulation run for one viscous time.
 The solid circles in figure~\ref{fig:LSB} represent runs where we performed GQL analyses.
@@
 In order to demonstrate the effects non-normality plays, figure~\ref{fig:pseudospectra} shows both the modal spectrum (in blue points) and the $\epsilon-$pseudospectrum for the supercritical and subcritical $\Reyn_i$.
-The contour value $\epsilon^{-1}$ represents the set of values
+The contour value $\log_{10} \epsilon^{-1}$ represents the set of values
 \begin{equation}
 \label{eqn:pseudospectrum}
-    \lVert (\sigma \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
+    \lVert (z \mathbf{I} - \Lop)^{-1} \rVert \geq \epsilon^{-1}
 \end{equation}
-for a given complex number $\sigma = \gamma + i\omega$ here representing the eigenvalue corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction).
+for a given complex number $z = \gamma + i\omega$.
+The spectrum {$\sigma$} are the eigenvalues corresponding to the temporal evolution of a linear solution, $e^{\sigma t}$ (see \cite{trefethen2005spectra} for an introduction).
 Equation~\ref{eqn:pseudospectrum} depends on the choice of norm; here we use the standard energy norm.
@@
 This linear effect is captured entirely by GQL (and QL, for that matter).
-What role it plays in the development of subcritical spiral turbulence remains an open question, though our results provide some interesting constraints.
+However, the role it plays in the development of subcritical spiral turbulence remains an open question, as the relevant linear operator $\Lop$ in the subcritical case is not that associated with circular Couette flow $\mathbf{U}$.
+Instead, any transient growth is generated by the saturated mean velocity profile.
 \begin{figure}
@@
     \end{minipage}
-    \caption{Spectra (blue points) and pseudospectra (contours) for $\Reyn_i = 900$ (left) and $\Reyn_i = 700$ (right) for a mode with $m=6$ and $k_z = 9.68$.
+    \caption{Spectra (blue points) and pseudospectra (contours of $\log_{10}\epsilon^{-1}$) for $\Reyn_i = 900$ (left) and $\Reyn_i = 700$ (right) for a mode with $m=6$ and $k_z = 9.68$.
 The latter is the z wavenumber at the critical $\Reyn_{i,crit} \simeq 778$.
@@
 In that work, the author proposed a potential explanation in terms of the linear instability of the non-linear solution: assume there is an instability of the mean $(m, k_z) = (0,0)$ at some wavenumber $k'$.
-If $k'$ is counted among the high modes, high-high $\to$ high interactions smoothly populate the high modes.
+If $k'$ is counted among the high modes, high-low $\to$ high interactions smoothly populate the high modes.
 However, as $\Lambda$ increases, it reaches a point at which the last unstable mode $k'$ becomes counted among the low modes.
@@
     \centering
-    \begin{minipage}{0.45\textwidth}
-        \centering
-        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
-    \end{minipage}
-    \begin{minipage}{0.45\textwidth}
-        \centering
+    %begin{minipage}{0.45\textwidth}
+    %    \centering
+        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}\\
+    %\end{minipage}
+    %\begin{minipage}{0.45\textwidth}
+    %    \centering
         \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
-    \end{minipage}
+    %\end{minipage}
     \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$.
-$\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary.
-$\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary.
-Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$.
-At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations.
+$\Reyn_i = 900$ (upper panel) is \emph{above} the linear stability boundary.
+$\Reyn_i = 700$ (lower panel), continued by lowering $\Reyn_i$ from previous simulations, is in the hysteretic region \emph{below} the stability boundary.
+Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (upper), $t/\tau_\nu \simeq 3$ (lower), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$.
+At higher $\Reyn_i$, all of the reduced models except $\Lambda = 60$ begin high amplitude oscillations.
 At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state.
@@
 The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.
-The figure shows only the lowest 100 modes in each direction, though in the simulations $k_{max} = m_{max} = 512$.
-}
+The figure shows only the lowest 100 modes in each direction, though in the simulations $|k_{max}| = 256$ and  $m_{max} = 511$.}
     \label{fig:rei900_spectra}
@@
 \cite{2009PhRvE..80d6315M}).
-At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating.
+At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating flows.
 As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT).
@@
 Prior work has focused on the utility of GQL in presenting a more accurate \emph{statistical} picture of turbulent flows than QL \cite{2016PhRvL.116u4501M, 2017JFM...810..412T,2018RSPSA.47480422T,2019Kellam}.
-This focus is particularly important in light of its potential as as means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realisations \cite{2022arXiv220505513M}.
+This focus is particularly important in light of its potential as a means of improving the accuracy of direct statistical simulation (DSS), a method for studying turbulence by computing its statistics directly, rather than individual realisations \cite{2022arXiv220505513M}.
 We have shown here that it is possible to use GQL as a means to investigate the non-linear interactions of greatest importance for maintaining the intricate patterns manifested in Taylor-Couette flow.
